    %! Author = Omar Iskandarani
    %! Title = VAM: Explaining the Universe in Analogies
    %! Date = xx-xx-2025
    %! Affiliation = Independent Researcher, Groningen, The Netherlands
    %! License = © 2025 Omar Iskandarani. All rights reserved. This manuscript is made available for academic reading and citation only. No republication, redistribution, or derivative works are permitted without explicit written permission from the author. Contact: info@omariskandarani.com
    %! ORCID = 0009-0006-1686-3961
    %! DOI = 10.5281/zenodo.xxxxxxx






% === Metadata ===
    \newcommand{\papertitle}{Knot-Based Particle Logic for Quantum Gates}
    \newcommand{\paperdoi}{10.5281/zenodo.xxxxxxxx}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,graphicx,cite}
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}

    \documentclass[a4paper,12pt]{article}
    \usepackage{../vamstyle}

    \usepackage{import}
    \usepackage{subfiles}

    \usepackage{caption}


    \input{../vamstyle.sty}
    \input{../vam_appendix_setup.sty}
    \begin{document}
\begin{abstract}
This brief explores the use of particle-topology analogs in the Vortex Æther Model (VAM) to construct physically realizable quantum gates. Knot-based particles such as the trefoil (electron), Hopf link (W boson), and Solomon link (e⁻/e⁺) are repurposed as topological logic units. Their genus, chirality, and mutual linkability determine allowable gate operations, entanglement fidelity, and reconnection behavior. This formulation proposes a scalable, geometric alternative to conventional circuit gate architectures.
\end{abstract}


\section*{1. Fermions as Bits, Bosons as Gates}
Trefoils encode spin-½ matter bits; Hopf and Solomon links act as vector gate operators, controlling transitions via knot reconnection. W bosons as Hopf links can flip chirality between adjacent trefoils \cite{Iskandarani2025,Penrose2004}.

\section*{2. Gate Topologies and Allowed Operations}
Logic operations (X, Z, CNOT) arise from linking and unlinking configurations. For example, a Hopf-linked trefoil pair undergoing a reconnection acts as a CNOT, while a single Solomon link can mediate a swap via chirality reversal \cite{Berloff2014,Hall2016}.

\section*{3. Topological Protection}
Because operations depend on conserved quantities (link number, genus), gates resist noise-induced errors. Unlike voltage-gated logic, VAM gates rely on swirl vector continuity and curvature invariants \cite{Kauffman1991}.

\section*{Conclusion}
A topologically derived logic from particle-knot analogs provides a geometric quantum computation language with built-in symmetry and error suppression.

\bibliographystyle{unsrt}
\bibliography{knot_logic}
\end{document}
