\section*{Helicity, Chirality, and Knot Topology (Writhe + Twist)}

The helicity of a vortex knot is a topological invariant closely related to the knot's chirality. In fluid mechanics, the total helicity $H$ of a closed vortex loop can be decomposed into contributions from the knot's writhe (Wr) and twist (Tw) -- essentially, the geometry of the loop's centerline and the twisting of vorticity around it. In fact, for a single knotted flux tube (or vortex filament), the Călugăreanu-White formula gives the linking number as $\text{Link} = \text{Wr} + \text{Tw}$, and the helicity is proportional to this sum~\cite{knot_theroy_in_fluid}. For example, in a magnetic flux tube of flux $\Phi$, one finds $H = (\text{Wr} + \text{Tw}),\Phi^2$\cite{knot_theroy_in_fluid}. By analogy, a vortex knot's helicity is determined by $W+T$, the sum of its writhe (how coiled or knotted its centerline is in space) and twist (internal twisting of vorticity along the tube)\cite{knot_theroy_in_fluid}.

\begin{itemize}
\item Chiral knots -- those distinguishable from their mirror images -- generally have nonzero $W+T$, endowing them with a net helicity (a preferred handedness of circulation in the \ae ther). A prime example is the trefoil knot, which is chiral and would carry a nonzero helicity in one orientation (and opposite helicity in the mirror orientation). These chiral vortex knots inject helicity flux into the surrounding \ae ther; $\mathbf{v}\cdot\boldsymbol{\omega}\neq 0$ in their vicinity, which, by Eq.(\ref{eq:helicity-time}), slows local time flow and creates a vortex-induced gravity well.

\item Achiral knots, by contrast, are symmetric under mirror reflection and thus carry \textit{vanishing net helicity}. The classic example is the \textit{figure-eight knot}, which is an amphichiral knot (identical to its mirror image). For such a structure, the contributions of writhe and twist cancel out to give $W+T \approx 0$. In essence, the figure-eight vortex's loops twist one way as much as the other, yielding no overall helicity in the \ae ther. This has profound dynamical implications: with $H \approx 0$, an achiral vortex does not induce the usual swirl gravity or time-dilation effects. The \ae ther flow around it carries no net helicity flux to slow clocks or produce a persistent low-pressure well. One can say the figure-eight spins both ways'' in balance, generating \textit{no screw-like time threading}. In terms of Eq.(\ref{eq:helicity-time}), for an ideal achiral knot $\mathbf{v}\cdot\boldsymbol{\omega}\to 0$, so the proper time increment $d\tau$ essentially equals the background time increment $dt$ -- no significant dilation. Equivalently, the chronometric ratio $d\tau/dN$ tends to 1 for achiral knots, where $N$ is the uniform \ae ther time. This corresponds to $d\tau/d\mathcal{N} \to 1$'' as the exclusion criterion: if a structure's proper time advances nearly unimpeded (equal to absolute time), it is not embedded in any gravitational potential well.
\end{itemize}

In the full VAM time-dilation formula, achiral knots effectively remove the helicity-dependent terms. For instance, the unified expression for local vs. absolute time includes subtractive contributions from swirl rotation and vorticity-induced mass. An achiral knot sets those terms to zero, yielding $d\tau/dN \approx 1$ (no slowing). Thus, the figure-eight or any achiral topology would experience negligible vortex-induced time dilation -- its internal clock $\tau$ ticks almost at the same rate as the cosmic \ae ther time $N$, even if it were placed deep in the galaxy. This is in stark contrast to chiral matter knots, whose $\tau$ can be substantially slowed by the galactic swirl field (e.g., near massive cores or in strong rotation).
