%! Author = omar.iskandarani
%! Date = 6/2/2025

To the Editors,

\textit{Foundations of Physics}


I am pleased to submit the enclosed manuscript entitled \grqq Swirl Clocks and Vorticity-Induced Gravity: A Fluid-Dynamical Framework for Time Dilation and Gravitation\textquotedblright for consideration. This paper proposes a reinterpretation of relativistic time dilation effects using a structured vortex æther model grounded in classical fluid dynamics.


Building upon the historical notion of a superfluid vacuum, this work introduces a physically intuitive mechanism for both special and general relativistic phenomena by modeling elementary particles as rotating vortex nodes in a flat, inviscid æther. Time is defined intrinsically by vortex core rotation, and gravitational effects are recovered through æther inflow velocity gradients. The resulting model reproduces standard relativistic time dilation formulas without invoking curved spacetime, and suggests experimentally testable deviations in regimes such as rotating superfluids and nanostructured vortex systems.


The paper aligns with the journal's mission to explore foundational theoretical physics, offering an alternative mechanistic picture with implications for gravitation, cosmology, and quantum timekeeping. I believe this submission will be of interest to readers engaged in emergent gravity theories, fluid analogues of spacetime, and reinterpretations of general relativity.


Thank you for considering this manuscript. I look forward to your response.


Sincerely,

Omar Iskandarani

Independent Researcher

Groningen, The Netherlands

[Email]

[ORCID]