%! Author = Omar Iskandarani
%! Date = 2025-06-13

% === Metadata ===
\newcommand{\appendixtitle}{Unified Time Dilation Formula in the Vortex \AE{}ther Model}
\newcommand{\appendixauthor}{Omar Iskandarani}
\newcommand{\appendixaffil}{Independent Researcher, Groningen, The Netherlands}
\newcommand{\appendixdoi}{10.5281/zenodo.15566319}
\newcommand{\appendixorcid}{0009-0006-1686-3961}

\ifdefined\standalonechapter\else
  % Standalone mode
  \documentclass[12pt]{article}
  \usepackage[a4paper, margin=2cm]{geometry}
  \usepackage{ifthen} % we can use it safely now
  \usepackage{import}
  \usepackage{subfiles}
  \usepackage{hyperref}
  \usepackage{graphicx}
  \usepackage{amsmath, amssymb, physics}
  \usepackage{siunitx}
  \usepackage{tikz}
  \usepackage{booktabs}
  \usepackage{caption}
  \usepackage{array, tabularx}
  \usepackage{listings}
  \usepackage{bookmark}
  \usepackage{newtxtext,newtxmath}
  \usepackage[scaled=0.95]{inconsolata}
  \usepackage{mathrsfs}
  \input{../vam_appendix_setup.sty}
  \begin{document}

  % === Title page ===
  \titlepageOpen

\begin{abstract}
We present a unified time dilation formula derived from the Vortex \AE{}ther Model (VAM), a fluid-dynamic reformulation of gravitation and mass-energy interactions. Unlike General Relativity, where mass and curvature govern clock rates, VAM attributes gravitational phenomena to quantized vorticity, æther circulation, and swirl-induced pressure gradients. The proposed equation replaces the Schwarzschild and Kerr metric terms with vortex core tangential velocities, swirl angular frequencies, and an effective mass derived from exponentially decaying æther density. A hybridization mechanism smoothly interpolates between vortex-scale gravity and classical Newtonian coupling at macroscopic distances. The final expression captures six physical effects within one coherent framework: (1) vortex-induced mass generation via circulation and helicity, (2) bubble-like volume expansion due to internal irrotational flow, (3) acceleration of this flow under compression, (4) thermal-like energy response from swirl speedup, (5) relativistic time dilation from æther puncture during motion, and (6) swirl-based core-local time. The result is a mathematically robust, numerically testable model that unifies quantum vortex dynamics with gravitational time effects and remains non-singular across all radial domains.
\end{abstract}

  \titlepageClose
\fi

% ============= Begin of content ============
\section{\appendixtitle}


\section*{Introduction}
In General Relativity (GR), time dilation is sourced by mass and rotation via the Schwarzschild and Kerr metrics, respectively. The Vortex \ae{}ther Model (VAM) replaces these with swirl-induced vorticity, æther density gradients, and circulation fields. We derive and present a unified time dilation expression that incorporates:

\begin{itemize}
    \item Vortex-induced mass formation,
    \item Frame-dragging via swirl angular velocity,
    \item Bubble swelling dynamics mimicking thermodynamic gas laws,
    \item Exponential decay of æther vorticity with distance,
    \item Hybridization to classical GR at large $r$.
\end{itemize}

\subsection{Final VAM Time Dilation Equation}
The unified time dilation equation in VAM becomes:

\begin{equation}
\boxed{
t_\text{adjusted} = \Delta t \cdot \sqrt{
1
- \frac{2 G_{\text{hybrid}}(r) M_{\text{hybrid}}(r)}{r c^2}
- \frac{C_e^2}{c^2} e^{-r/r_c}
- \frac{C_e^2}{r_c^2 c^2} e^{-r/r_c}
}
}
\label{eq:final_vam_td}
\end{equation}

\subsection{Explanation of Terms}
\begin{enumerate}
    \item The first term \(\frac{2 G_{\text{hybrid}} M_{\text{hybrid}}}{r c^2}\) generalizes Newtonian gravity using hybrid interpolation.
    \item The second term \(\frac{C_e^2}{c^2} e^{-r/r_c}\) accounts for tangential swirl effects.
    \item The third term \(\frac{C_e^2}{r_c^2 c^2} e^{-r/r_c}\) models rotational energy frame-dragging.
\end{enumerate}


\begin{table}[h]
      \centering
      \begin{tabular}{llll}
          \toprule
          \textbf{Symbol} & \textbf{Meaning} & \textbf{Value / Expression} & \textbf{Units} \\
          \midrule
        $G_{\text{hybrid}}(r)$ & Hybrid gravitational constant (VAM/GR) & $\mu(r) G_{\text{swirl}} + (1 - \mu(r)) G$ & $\text{m}^3\,\text{kg}^{-1}\,\text{s}^{-2}$ \\
        $\mu(r)$ & Vortex-to-classical transition function & $e^{-r^2 / R_0^2}$, $R_0 = 1.0 \times 10^{-12}\,\text{m}$ & unitless \\
        $G$ & Newtonian gravitational constant & $6.67430 \times 10^{-11}$ & $\text{m}^3\,\text{kg}^{-1}\,\text{s}^{-2}$ \\
        $G_{\text{swirl}}$ & Swirl-induced gravitational constant & $\dfrac{C_e c^5 t_p^2}{2 F_{\max} r_c^2}$ & $\text{m}^3\,\text{kg}^{-1}\,\text{s}^{-2}$ \\
        $M_{\text{hybrid}}(r)$ & Hybrid effective mass & $\mu(r) M_\text{eff}^\text{VAM}(r) + (1 - \mu(r)) M$ & $\text{kg}$ \\
        $M_\text{eff}^\text{VAM}(r)$ & Vortex effective mass & $4\pi \rho_\text{\ae} r_c^3 \left[ 2 - (2 + \frac{r}{r_c}) e^{-r/r_c} \right]$ & $\text{kg}$ \\
        $\rho_\text{\ae}$ & Æther density & $3.89343583 \times 10^{18}$ & $\text{kg}\cdot\text{m}^{-3}$ \\
        $r_c$ & Core radius (Coulomb scale) & $1.40897017 \times 10^{-15}$ & $\text{m}$ \\
        $C_e$ & Core tangential velocity & $1.09384563 \times 10^{6}$ & $\text{m}\cdot\text{s}^{-1}$ \\
        $t_p$ & Planck time & $5.391247 \times 10^{-44}$ & $\text{s}$ \\
        $F_{\max}$ & Maximum force & $29.053507$ & $\text{N}$ \\
        $\left(\frac{C_e}{r_c}\right)^2$ & Squared swirl angular frequency ($\Omega^2$) & $6.02367430 \times 10^{42}$ & $\text{s}^{-2}$ \\
        $c$ & Speed of light & $2.99792458 \times 10^8$ & $\text{m}\cdot\text{s}^{-1}$ \\
          \bottomrule
      \end{tabular}
    \caption{Key symbols and constants in the VAM time dilation equation.}
      \label{tab:time_dilation_symbols}
  \end{table}

\subsection{Hybridization of Gravitational Coupling}

To reconcile predictions for macroscopic systems, we define:
\[
    \mu(r) = \exp\left(-\frac{r^2}{R_0^2}\right), \quad R_0 \sim 10^{-12} \, \mathrm{m}
\]
\begin{align*}
    G_{\text{hybrid}}(r) &= \mu(r) \, G_{\text{swirl}} + (1 - \mu(r)) \, G \\
    M_{\text{hybrid}}(r) &= \mu(r) \, M_{\text{eff}}^\text{VAM}(r) + (1 - \mu(r)) \, M
\end{align*}

\subsection{Effective VAM Mass}
Assuming an exponentially decaying æther density:
\[
\rho_\text{\ae}(r) = \rho_0 e^{-r / r_c}
\]
The effective mass becomes:
\[
M_\text{eff}^\text{VAM}(r) = 4\pi \rho_0 r_c^3 \left(2 - \left(2 + \frac{r}{r_c} \right) e^{-r/r_c} \right)
\]

\subsection{Constants and Variables}

\begin{table}
    \centering
    \begin{tabular}{llll}
        \toprule
        \textbf{Symbol} & \textbf{Meaning} & \textbf{Description} & \textbf{Value (if constant)} \\
        \midrule
        $\Delta t$ & Reference time & Clock far from gravitating body & -- \\
        $t_\text{adjusted}$ & Local time & Time experienced near the vortex structure & -- \\
        $r$ & Radial coordinate & Distance from the vortex core & m \\
        $r_c$ & Vortex core radius & Characteristic decay scale & $1.40897017 \times 10^{-15}$ m \\
        $C_e$ & Vortex tangential velocity & Maximal edge swirl velocity & $1.09384563 \times 10^6$ m/s \\
        $\rho_\text{\ae}$ & Æther density & Fluid density of the æther & $\sim 3.89 \times 10^{18}$ kg/m$^3$ \\
        $c$ & Speed of light & Vacuum light speed & $2.99792458 \times 10^8$ m/s \\
        $G$ & Newton's constant & Classical gravity & $6.67430 \times 10^{-11}$ m$^3$/kg/s$^2$ \\
        $F_{\text{max}}$ & Max force & From Planck-scale dynamics & $29.053507$ N \\
        $t_p$ & Planck time & Quantum gravity scale & $5.391247 \times 10^{-44}$ s \\
        $G_\text{swirl}$ & Vortex gravity coupling & $C_e c^5 t_p^2 / (2 F_{\text{max}} r_c^2)$ & -- \\
        $M$ & Macroscopic mass & Classical object mass (e.g., proton mass) & $1.67262192 \times 10^{-27}$ kg \\
        $M_{\text{eff}}^\text{VAM}(r)$ & VAM mass & Mass from vorticity energy & derived \\
        $M_{\text{hybrid}}(r)$ & Hybrid mass & Smooth transition between VAM and GR & -- \\
        $G_{\text{hybrid}}(r)$ & Hybrid gravity constant & Smooth transition between $G$ and $G_\text{swirl}$ & -- \\
        $\mu(r)$ & Hybrid blending function & $\mu(r) = \exp\left(-\frac{r^2}{R_0^2}\right),\ R_0 \sim 10^{-12}$ m & dimensionless \\
        $e^{-r/r_c}$ & Vorticity decay & Exponential suppression term & -- \\
        \bottomrule
    \end{tabular}
    \caption{Explanation of variables in Equation~\ref{eq:final_vam_td}.}
    \label{tab:time_dilation_variables}
\end{table}



\section*{Conclusion}
This equation synthesizes all prior VAM elements: vortex helicity, bubble boundaries, circulation-induced gravity, and exponential suppression of short-range fields. It remains finite, matches classical predictions at macroscopic scales, and enables numerical probing at quantum scales.

% ============== End of content =============

% === Bibliography (only for standalone) ===
\ifdefined\standalonechapter
  % Being imported from main.tex — do nothing
\else
    \bibliographystyle{unsrt}
    \bibliography{../../references}
    \end{document}
\fi