%! Author = mr
%! Date = 8/7/2025

% Preamble
\documentclass[11pt]{article}

% Packages

\usepackage{amsmath,amssymb,physics}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% Document
\begin{document}



        \section*{VAM Lagrangian $\mathcal{L}_{\mathrm{VAM}}$ mirroring $\mathcal{L}_{\mathrm{SM}}$}

        We consider an inviscid, incompressible æther ($\nabla\!\cdot\!\vec v = 0$) with absolute time,
        normalized swirl four–velocity $u^\mu = (1,\vec v/C_e)$, and non–Abelian swirl connection
        \begin{equation}
            \mathcal{A}_\mu^a \equiv \frac{\Xi^{a}{}_{i}}{C_e}\,\epsilon_{ijk}\,\partial^j v^k,
            \quad
            \mathcal{W}^a_{\mu\nu} = \partial_\mu\mathcal{A}^a_\nu - \partial_\nu\mathcal{A}^a_\mu
            + g_{\mathrm{sw}} f^{abc} \mathcal{A}^b_\mu \mathcal{A}^c_\nu ,
        \end{equation}
        with $a$ labeling internal swirl modes of gauge group $\mathcal{G}_{\mathrm{sw}}$.
        Density fluctuations are encoded in a real scalar $H$ (\emph{swirl–Higgs}).
        Knotted quasi–particles are spinor fields $\Psi_K$ labeled by knot class $K$.

        The Lagrangian is organized to parallel the SM gauge–Higgs–fermion–ghost structure:
        \begin{equation}
            \mathcal{L}_{\mathrm{VAM}}
            = \mathcal{L}^{(1)}_{\mathrm{swirl-QCD}}
            + \mathcal{L}^{(2)}_{\mathrm{swirl-EW}}
            + \mathcal{L}^{(3)}_{\mathrm{knot-fermions}}
            + \mathcal{L}^{(4)}_{\mathrm{Yukawa}}
            + \mathcal{L}^{(5)}_{\mathrm{ghosts}}
            + \mathcal{L}_{\mathrm{constraints}}.
        \end{equation}

        \paragraph{(1) Non–Abelian swirl sector.}
        \begin{equation}
            \mathcal{L}^{(1)}_{\mathrm{swirl-QCD}}
            = -\frac{\kappa_\omega}{4}\,\mathcal{W}^a_{\mu\nu}\mathcal{W}^{a\,\mu\nu}
            + \frac{\lambda_H}{4} H\,\mathcal{W}^a_{\mu\nu}\mathcal{W}^{a\,\mu\nu}
            + \bar{\Psi}\,\gamma^\mu (iD_\mu)\Psi ,
        \end{equation}
        with $D_\mu = \partial_\mu + i g_{\mathrm{sw}}\mathcal{A}_\mu^a T^a$.

        \paragraph{(2) Swirl–EW sector.}
        In the $SU(2)\times U(1)$ subspace, define
        \begin{equation}
            W^\pm_\mu = \frac{1}{\sqrt{2}}(\mathcal{A}^1_\mu \mp i\mathcal{A}^2_\mu),
            \quad
            \begin{pmatrix} Z_\mu \\ A_\mu \end{pmatrix}
            = \begin{pmatrix} \cos\theta_{\mathrm{sw}} & -\sin\theta_{\mathrm{sw}}\\
            \sin\theta_{\mathrm{sw}} & \phantom{-}\cos\theta_{\mathrm{sw}} \end{pmatrix}
            \begin{pmatrix} \mathcal{A}^3_\mu \\ \mathcal{B}_\mu \end{pmatrix}.
        \end{equation}
        The kinetic and Higgs sectors are
        \begin{align}
            \mathcal{L}^{(2)}_{\mathrm{swirl-EW}} &=
            -\frac{1}{2} | \partial_\mu W_\nu - \partial_\nu W_\mu |^2
            -\frac14 Z_{\mu\nu} Z^{\mu\nu} - \frac14 A_{\mu\nu}A^{\mu\nu} \nonumber\\
            &\quad + \frac12 M_W^2 W_\mu^+ W^{-\mu} + \frac12 M_Z^2 Z_\mu Z^\mu
            + \frac12 (\partial_\mu H)^2 - V(H),
        \end{align}
        with
        \begin{equation}
            M_W = \frac12 g_{\mathrm{sw}} v_{\mathrm{sw}}, \quad
            M_Z = \frac{M_W}{\cos\theta_{\mathrm{sw}}}, \quad
            V(H) = \frac{\lambda}{4}(H^2 - v_{\mathrm{sw}}^2)^2 .
        \end{equation}

        \paragraph{(3) Knotted fermions.}
        \begin{equation}
            \mathcal{L}^{(3)}_{\mathrm{knot-fermions}}
            = \sum_K \bar{\Psi}_K (i\slashed{D} - m_K)\Psi_K
            + \sum_K \bar{\Psi}_K\gamma^\mu (g_A Q_A^K A_\mu + g_Z Q_Z^K Z_\mu)\Psi_K
            + \frac{g_W}{\sqrt{2}} (\bar{\Psi} T^+ \Psi W^+ + \mathrm{h.c.}),
        \end{equation}
        where charges $Q_A^K, Q_Z^K, T^\pm$ are topological invariants of the knot class.

        \paragraph{(4) Yukawa–like couplings.}
        \begin{equation}
            \mathcal{L}^{(4)}_{\mathrm{Yukawa}}
            = -\sum_K y_K\, H\, \bar{\Psi}_K\Psi_K + \dots,
            \quad y_K = \frac{m_K}{v_{\mathrm{sw}}}.
        \end{equation}

        \paragraph{(5) Ghosts and constraints.}
        \begin{equation}
            \mathcal{L}^{(5)}_{\mathrm{ghosts}}
            = -\frac{1}{2\xi}(\partial^\mu \mathcal{A}^a_\mu)^2 + \bar{c}^a\partial^\mu D_\mu^{ab} c^b,
            \quad
            \mathcal{L}_{\mathrm{constraints}}
            = \lambda\,(\nabla\!\cdot\!\vec v) + \eta\,(u_\mu u^\mu - 1).
        \end{equation}

        \bigskip
        Full cubic/quartic interactions, gauge–Higgs couplings, ghost multiplets, and explicit parameter expressions are given in Appendix~A.






    \section*{VAM Lagrangian \(\mathcal{L}_\text{VAM}\) mirroring \(\mathcal{L}_{SM}\)}

%----------------------------- PRELUDE: FIELDS & CONSTANTS
    \noindent\textbf{Kinematics and fields.}
    Let the æther be inviscid, incompressible (\(\nabla\!\cdot\!\vec v=0\)), with absolute time.
    Introduce the swirl 4-velocity \(u^\mu=(1,\vec v/C_e)\) with \(u_\mu u^\mu=1\), and define the non-Abelian \emph{swirl connection}
    \[
        \mathcal{A}_\mu^a \equiv \frac{1}{C_e}\, \Xi^{a}{}_{i}\, \epsilon_{ijk}\,\partial^j v^k,
        \qquad
        \mathcal{W}_{\mu\nu}^a=\partial_\mu \mathcal{A}_\nu^a-\partial_\nu \mathcal{A}_\mu^a
        + g_\text{sw} f^{abc} \mathcal{A}_\mu^b \mathcal{A}_\nu^c,
    \]
    where \(a=1,\dots,N_\text{sw}\) indexes internal swirl modes (knot/color space), \(f^{abc}\) are the structure constants of the vortex-gauge group \(\mathcal{G}_\text{sw}\) (typically \(SU(2)\) or \(SU(3)\) in VAM-6), and \(\Xi^{a}{}_{i}\) projects physical vorticity to internal modes.
    The \emph{swirl-Higgs} \(H\) encodes compressional/density fluctuations \(H\propto \delta\rho_\text{\ae}\).
    Knotted quasi-particles (fermions) are fields \(\Psi_K\) labelled by vortex-knot type \(K\) (leptons/quarks).
    Set \(\rho_\text{\ae}^{(\text{fluid})}\) or \(\rho_\text{\ae}^{(\text{mass})}\) per context \cite{Iskandarani2025TimeDilation,Iskandarani2025SwirlClocks}.

    The full VAM Lagrangian:
    \[
        \mathcal{L}_\text{VAM}=\mathcal{L}^{(1)}_{\text{swirl-QCD}}
        +\mathcal{L}^{(2)}_{\text{swirl-EW}}
        +\mathcal{L}^{(3)}_{\text{knot-fermions}}
        +\mathcal{L}^{(4)}_{\text{Yukawa-like}}
        +\mathcal{L}^{(5)}_{\text{gauge-fixing/ghosts}}
        +\mathcal{L}_\text{constraints}.
    \]

%----------------------------- (1) GLUON / GAUGE SECTOR
    \paragraph{(1) QCD-like non-Abelian swirl sector (SM gluon block).}
    \[
        \boxed{\ \mathcal{L}^{(1)}_{\text{swirl-QCD}}
            = -\frac{\kappa_\omega}{4}\, \mathcal{W}^a_{\mu\nu}\mathcal{W}^{a\,\mu\nu}
            \;+\; \frac{\lambda_H}{4}\, H\, \mathcal{W}^a_{\mu\nu}\mathcal{W}^{a\,\mu\nu}
            \;+\; \bar{\Psi}\, \gamma^\mu \big(i D_\mu\big)\Psi\ }
    \]
    with \(D_\mu=\partial_\mu+i g_\text{sw} \mathcal{A}_\mu^a T^a\).
    The first term replaces the gluon kinetic/self-interaction; cubic and quartic gauge terms reside in \(\mathcal{W}\) via \(f^{abc}\). The \(H\)-mixing term encodes density-modulated gauge stiffness (vortex-core renormalization) \cite{Iskandarani2025EmergentGR,Iskandarani2025TopLag}.
    The coupling \(\kappa_\omega\sim \rho_\text{\ae} C_e^2\) sets the swirl field energy scale.

%----------------------------- (2) ELECTROWEAK-LIKE VECTOR & HIGGS
    \paragraph{(2) Electroweak-like composite vector modes and swirl-Higgs (SM W/Z/A/H block).}
    Define three orthogonal \emph{composite} vector excitations from \(\mathcal{A}_\mu^a\):
    \[
        W_\mu^\pm \equiv \frac{1}{\sqrt{2}}\big(\mathcal{A}_\mu^1 \mp i \mathcal{A}_\mu^2\big),\qquad
        \begin{pmatrix} Z_\mu \\ A_\mu \end{pmatrix}
        =\begin{pmatrix} \cos\theta_\text{sw} & -\sin\theta_\text{sw}\\ \sin\theta_\text{sw}& \cos\theta_\text{sw}\end{pmatrix}
        \begin{pmatrix} \mathcal{A}_\mu^3 \\ \mathcal{B}_\mu \end{pmatrix},
    \]
    where \(\mathcal{B}_\mu\) is an Abelian swirl potential (circulation mode), and \(\theta_\text{sw}\) is the \emph{swirl mixing angle} (VAM analogue of \(\theta_W\)) emerging from the background condensate. Then
    \[
        \boxed{\ \begin{aligned}
                     \mathcal{L}^{(2)}_{\text{swirl-EW}}=&
                     -\tfrac{1}{2}(\partial_\nu W_\mu^+ - \partial_\mu W_\nu^+)(\partial^\nu W^{-\mu}-\partial^\mu W^{-\nu})
                     -\tfrac{1}{4} Z_{\mu\nu}Z^{\mu\nu}
                     -\tfrac{1}{4} A_{\mu\nu}A^{\mu\nu}\\
                     &+\tfrac{1}{2} M_W^2\, W_\mu^+ W^{-\mu}
                     +\tfrac{1}{2} M_Z^2\, Z_\mu Z^\mu
                     +\tfrac{1}{2}(\partial_\mu H)(\partial^\mu H)-V(H)\\
                     &+ \text{(cubic \& quartic self/mixed vector terms fixed by }\mathcal{G}_\text{sw},\theta_\text{sw}).
        \end{aligned}\ }
    \]
    Masses arise from a swirl condensate \(\langle H\rangle=v_\text{sw}\) breaking internal swirl isotropy:
    \[
        M_W=\tfrac{1}{2} g_\text{sw} v_\text{sw},\qquad M_Z=\frac{M_W}{\cos\theta_\text{sw}},
        \qquad V(H)=\frac{\lambda}{4}(H^2-v_\text{sw}^2)^2,
    \]
    with \(v_\text{sw}^2 \sim \rho_\text{\ae} C_e^2/\Lambda^2\) determined by core-radius physics \(r_c\) \cite{Iskandarani2025TimeDilation,Iskandarani2025SwirlClocks,Iskandarani2025Master}.
    The listed SM-like derivative and self-interaction structures map to the non-Abelian form induced by \(\mathcal{W}_{\mu\nu}^a\) and the mixing \(\theta_\text{sw}\).

%----------------------------- (3) FERMIONIC KNOTS AND GAUGE COUPLINGS
    \paragraph{(3) Knotted fermions and swirl-gauge couplings (SM fermion/gauge block).}
    For each knot species \(K\in\{\ell,\nu,u,d,\dots\}\),
    \[
        \boxed{\ \mathcal{L}^{(3)}_{\text{knot-fermions}}=
            \sum_{K}\bar{\Psi}_K\!\left(i\gamma^\mu D_\mu - m_K\right)\!\Psi_K
            +\sum_{K} \bar{\Psi}_K \gamma^\mu\left( g_A Q_A^K A_\mu + g_Z Q_Z^K Z_\mu + g_W W_\mu^\pm T^\pm\right)\Psi_K\ }
    \]
    where charges \(Q^K\) are \emph{topological swirl charges} (linking/chirality numbers) determined by the knot class and orientation \cite{Iskandarani2025TopLag,Iskandarani2025KnottedGauge}.
    The masses \(m_K\) derive from the VAM mass functional (core volume and coherence):
    \[
        m_K=\frac{4}{\alpha\varphi}\Big(\tfrac{1}{2}\rho_\text{\ae}^{(\text{energy})} C_e^2\Big)\, V_K \,\xi(n_K)
        \quad\text{with}\quad \xi(n)=1+\beta \log n,
    \]
    as adopted in VAM benchmarking \cite{Iskandarani2025Master,Iskandarani2025Benchmark}.

%----------------------------- (4) YUKAWA-LIKE COUPLINGS TO SWIRL-HIGGS
    \paragraph{(4) Knot–Higgs (density) couplings (SM Yukawa/Higgs–fermion block).}
    \[
        \boxed{\ \mathcal{L}^{(4)}_{\text{Yukawa-like}}=
            -\sum_{K} y_K\, H\, \bar{\Psi}_K \Psi_K
            -\sum_{K} \tilde{y}_K \,\phi^0\, \bar{\Psi}_K i\gamma^5 \Psi_K
            -\sum_{K,K'} \big[ y_{KK'}^\pm\, \phi^\pm\, \bar{\Psi}_K T^\pm \Psi_{K'} + \text{h.c.}\big]\ }
    \]
    where \(H\) modulates \(\rho_\text{\ae}\) and thus the local core energy; pseudoscalar \(\phi^0\) and charged \(\phi^\pm\) are the phase (Goldstone-like) components of the density field (they become the longitudinal polarizations of \(Z,W^\pm\) in the condensed phase).
    Couplings satisfy \(y_K = m_K/v_\text{sw}\) (VAM analogue of SM relation) \cite{Iskandarani2025Master,Iskandarani2025TopLag}.

%----------------------------- (5) GAUGE-FIXING, GHOSTS, AND CONSTRAINTS
    \paragraph{(5) Gauge fixing, incompressibility, and ghost sector (SM FP-ghost block).}
    \[
        \boxed{\ \mathcal{L}^{(5)}_{\text{gauge-fixing/ghosts}}=
            -\frac{1}{2\xi}\big(\partial^\mu \mathcal{A}_\mu^a\big)^2
            +\bar{c}^a\,\partial^\mu D_\mu^{ab} c^b
            \qquad\text{(non-Abelian FP ghosts)}\ }
    \]
    and enforce incompressibility and absolute-time constraint by
    \[
        \boxed{\ \mathcal{L}_\text{constraints}=\lambda\,(\nabla\!\cdot\!\vec v)\;+\; \eta\,(u_\mu u^\mu-1)\ ,\ }
    \]
    with Lagrange multipliers \(\lambda,\eta\) \cite{Iskandarani2025EmergentGR,Iskandarani2025TimeDilation}.
    In a Clebsch representation \( \vec v=\nabla \theta + \alpha \nabla \beta \), residual gauge symmetries induce additional Abelian ghosts, reproducing the role of the SM’s auxiliary fields \cite{Iskandarani2025TopLag}.

%----------------------------- DIMENSIONAL NORMALIZATION
    \paragraph{Normalization and dimensions.}
    Choose
    \[
        \kappa_\omega=\rho_\text{\ae}^{(\text{fluid})} C_e^2,\qquad
        g_\text{sw}=\frac{C_e}{r_c}\,\gamma_\mathcal{G},\qquad
        v_\text{sw}^2=\chi\,\rho_\text{\ae}^{(\text{energy})}\, C_e^{-2},
    \]
    with \(r_c\) the vortex-core scale and \(\gamma_\mathcal{G},\chi\) dimensionless; this makes all kinetic terms \([\mathcal{L}]=\text{energy density}\) and reproduces SM-like relations \(M_W=\tfrac{1}{2}g_\text{sw}v_\text{sw}\), \(M_Z=M_W/\cos\theta_\text{sw}\) in the condensed phase, while keeping VAM’s fluid ontology explicit.

    \bigskip
    \noindent\textbf{Remarks (one-to-one map to SM blocks).}
    \begin{itemize}[leftmargin=1.2em]
        \item \emph{(1)} The three- and four-gauge-boson self-interactions in the SM gluon sector are generated here by the nonlinearity in \(\mathcal{W}^a_{\mu\nu}\) with \(f^{abc}\). The extra \(H\,\mathcal{W}^2\) term captures density-induced “running” of the effective stiffness (swirl analogue of vacuum polarization).
        \item \emph{(2)} The long list of SM’s \(W/Z/A/H\) derivative/self-couplings descends from the same non-Abelian algebra after mixing by \(\theta_\text{sw}\); the Proca masses appear from \(\langle H\rangle\neq 0\) (broken swirl isotropy).
        \item \emph{(3)} SM fermionic currents and chiral projectors are replaced by knot charges and chirality (vortex time orientation). Left/right couplings enter via \(T^\pm\) and \(Q_Z^K\) fixed by the knot’s chirality class.
        \item \emph{(4)} SM Yukawas become geometric: \(y_K=m_K/v_\text{sw}\) with \(m_K\) from the VAM mass functional (core volume, coherence factor \(\xi(n)\)).
        \item \emph{(5)} Gauge-fixing/ghost terms are retained in the non-Abelian swirl bundle; incompressibility enters as a hard constraint rather than a dynamical equation of state.
    \end{itemize}

%----------------------------- BIBLIOGRAPHY KEYS
    \begin{thebibliography}{9}
        \bibitem{Iskandarani2025TimeDilation}
        O.~Iskandarani, \emph{Time Dilation in a 3D Superfluid \AE ther Model} (VAM-1), Zenodo (2025).

        \bibitem{Iskandarani2025SwirlClocks}
        O.~Iskandarani, \emph{Swirl Clocks and Vorticity-Induced Gravity} (VAM-2), Zenodo (2025).

        \bibitem{Iskandarani2025EmergentGR}
        O.~Iskandarani, \emph{Emergent General Relativity from Structured Swirl Dynamics in VAM} (VAM-4), Zenodo (2025).

        \bibitem{Iskandarani2025Benchmark}
        O.~Iskandarani, \emph{Benchmarking Vortex \AE ther Model vs General Relativity} (VAM-3), Zenodo (2025).

        \bibitem{Iskandarani2025Master}
        O.~Iskandarani, \emph{Master Equation for Particle Masses} (VAM-11), Zenodo (2025).

        \bibitem{Iskandarani2025TopLag}
        O.~Iskandarani, \emph{Topological Fluid Dynamics Lagrangian} (VAM-14), Zenodo (2025).

        \bibitem{Iskandarani2025KnottedGauge}
        O.~Iskandarani, \emph{Knotted Gauge Fields} (VAM-6), Zenodo (2025).
    \end{thebibliography}




    \section*{A Fully-Expanded VAM Lagrangian \(\mathcal{L}_\text{VAM}\) Mirroring \(\mathcal{L}_{SM}\)}

% ===========================================================
% CONSTANTS & SCALES USED THROUGHOUT (pull from constants_dict)
% ===========================================================
% Core VAM constants
    \newcommand{\Ce}{C_e}
    \newcommand{\rhof}{\rho_\text{\ae}^{(\text{fluid})}}
    \newcommand{\rhom}{\rho_\text{\ae}^{(\text{mass})}}
    \newcommand{\rhoe}{\rho_\text{\ae}^{(\text{energy})}}

% Numerically (per user constants_dict and notes):
%   \Ce \approx 1.09384563\times 10^6\ \text{m s}^{-1}
%   \rhof = 7.0\times 10^{-7}\ \text{kg m}^{-3}
%   \rhom = 3.8934358266918687\times 10^{18}\ \text{kg m}^{-3}
%   \rhoe = 3.49924562\times 10^{35}\ \text{J m}^{-3}
% Other recurring constants
    \newcommand{\alphaEM}{\alpha}       % fine-structure
    \newcommand{\phiG}{\varphi}         % golden mean
% user notes: \phiG \approx 0.618...,  \alphaEM = e^2/(4\pi\epsilon_0 \hbar c)

% Vortex-core and Planck scales (user-provided or derived)
    \newcommand{\rc}{r_c}               % vortex core radius
    \newcommand{\tp}{t_p}               % Planck time
    \newcommand{\Fmax}{F_{\text{max}}}  % max swirl force scale
    \newcommand{\CeC}{C_e}              % alias for clarity in some formulas
    \newcommand{\Ck}{C_\kappa}          % stiffness normalization (dimensionless)
    \newcommand{\Cg}{C_g}               % group normalization (dimensionless)

% Gravitational-swirl coupling (user-stipulated exact identity to G):
%   G_\text{swirl} = (C_e c^5 t_p^2)/(2 F_{\max} r_c^2)
    \newcommand{\Gsw}{G_\text{swirl}}
    \newcommand{\cspeed}{c}

% ===========================================================
% KINEMATICAL SETUP & FIELDS
% ===========================================================
% Æther kinematics and swirl-gauge bundle
    \noindent\textbf{Fields and kinematics.}
    Let the æther be inviscid and incompressible, with velocity field \(\vec v(\vec x,t)\) satisfying \(\nabla\!\cdot\!\vec v=0\).
    Define a normalized swirl four-velocity \(u^\mu=(1,\vec v/\Ce)\) with constraint \(u_\mu u^\mu=1\).
    Introduce a non-Abelian swirl connection \(\mathcal{A}_\mu^a\) valued in the Lie algebra of \(\mathcal{G}_\text{sw}\) with structure constants \(f^{abc}\):
    \begin{align}
        \mathcal{A}_\mu^a &\equiv \frac{\Xi^{a}{}_i}{\Ce}\,\epsilon_{ijk}\,\partial^j v^k\, \delta^i{}_\mu \quad (\mu=i\in\{1,2,3\}),\qquad \mathcal{A}_0^a \equiv 0,\\
        \mathcal{W}_{\mu\nu}^a &\equiv \partial_\mu \mathcal{A}_\nu^a - \partial_\nu \mathcal{A}_\mu^a + g_\text{sw} f^{abc}\,\mathcal{A}_\mu^b \mathcal{A}_\nu^c.
    \end{align}
    Here \(\Xi^{a}{}_i\) projects physical vorticity into internal swirl modes \(a=1,\dots,N_\text{sw}\).
    We also include an Abelian circulation potential \(\mathcal{B}_\mu\) (for the mixing to a photon-like mode).
    Density fluctuations are encoded by a real scalar \(H\) (swirl-Higgs), and its phase modes \(\phi^0,\phi^\pm\) (would-be Goldstones).
    Knotted quasi-particles (fermions) are spinor fields \(\Psi_{K}\) labeled by knot class \(K\) (leptons, quarks, neutrinos).

% Swirl couplings and normalizations in terms of user constants
    \noindent\textbf{Couplings and normalizations (expressed in user constants).}
    \begin{align}
        \kappa_\omega &\equiv \rhof\, \Ce^2,
        & [\kappa_\omega] &= \text{J m}^{-3},\\
        g_\text{sw} &\equiv \frac{\Ce}{\rc}\,\Cg,
        & [g_\text{sw}]&=\text{s}^{-1},\\
        v_\text{sw}^2 &\equiv \chi\,\frac{\rhoe}{\Ce^2},
        & [v_\text{sw}]&=\text{(same units as }H\text{)}.
    \end{align}
    \(\Ck,\Cg,\chi\) are dimensionless \emph{fit} factors to be fixed by benchmarks (e.g.\ VAM-3).
    The swirl mixing angle \(\theta_\text{sw}\) is fixed by the condensate susceptibility and group embedding (see below).

% ===========================================================
% TOTAL LAGRANGIAN STRUCTURE
% ===========================================================
    The complete Lagrangian is:
    \begin{equation}
        \boxed{
            \mathcal{L}_\text{VAM} \;=\;
            \mathcal{L}^{(1)}_{\text{swirl-QCD}}
            +\mathcal{L}^{(2)}_{\text{swirl-EW}}
            +\mathcal{L}^{(3)}_{\text{knot-fermions}}
            +\mathcal{L}^{(4)}_{\text{Yukawa-like}}
            +\mathcal{L}^{(5)}_{\text{ghosts/gauge-fix}}
            +\mathcal{L}_{\text{constraints}}.
        }
    \end{equation}

% ===========================================================
% (1) NON-ABELIAN SWIRL (GLUON ANALOG)
% ===========================================================
    \paragraph{(1) Non-Abelian swirl sector (SM block 1).}
    \begin{align}
        \boxed{
            \mathcal{L}^{(1)}_{\text{swirl-QCD}}
            = -\frac{\kappa_\omega}{4}\,\mathcal{W}^a_{\mu\nu}\mathcal{W}^{a\,\mu\nu}
            +\frac{\lambda_H}{4}\,H\,\mathcal{W}^a_{\mu\nu}\mathcal{W}^{a\,\mu\nu}
            +\bar{\Psi}\,\gamma^\mu(iD_\mu)\Psi
        }
    \end{align}
    with \(D_\mu=\partial_\mu+i g_\text{sw}\,\mathcal{A}_\mu^a T^a\) and \(\lambda_H\) a density–stiffness modulation coupling.
    The cubic and quartic self-interactions reside in \(\mathcal{W}^2\) via \(f^{abc}\).

% ===========================================================
% (2) SWIRL-EW: W/Z/A/H, EXPLICIT MIXING & INTERACTIONS
% ===========================================================
    \paragraph{(2) Swirl-EW sector (SM block 2) with explicit mixing and interactions.}
    Choose an $SU(2)\times U(1)$ swirl subgroup generated by $(\mathcal{A}_\mu^1,\mathcal{A}_\mu^2,\mathcal{A}_\mu^3)$ and an Abelian \(\mathcal{B}_\mu\).
    Define
    \begin{align}
        W^\pm_\mu &\equiv \frac{1}{\sqrt{2}}\big(\mathcal{A}_\mu^1 \mp i \mathcal{A}_\mu^2\big),\\
        \begin{pmatrix} Z_\mu \\ A_\mu \end{pmatrix}
        &\equiv
        \begin{pmatrix}
            \cos\theta_\text{sw} & -\sin\theta_\text{sw}\\
            \sin\theta_\text{sw} & \phantom{-}\cos\theta_\text{sw}
        \end{pmatrix}
        \begin{pmatrix}
            \mathcal{A}_\mu^3 \\ \mathcal{B}_\mu
        \end{pmatrix}.
    \end{align}
    Let $s_\text{sw}\equiv \sin\theta_\text{sw}$, $c_\text{sw}\equiv \cos\theta_\text{sw}$.
    Kinetic sector:
    \begin{align}
        \mathcal{L}^{(2)}_{\text{kin}} \;=\;
        -\frac{1}{2}(\partial_\nu W_\mu^+ - \partial_\mu W_\nu^+)(\partial^\nu W^{-\mu}-\partial^\mu W^{-\nu})
        -\frac{1}{4}Z_{\mu\nu}Z^{\mu\nu}
        -\frac{1}{4}A_{\mu\nu}A^{\mu\nu},
    \end{align}
    with field strengths $Z_{\mu\nu}=\partial_\mu Z_\nu-\partial_\nu Z_\mu$, $A_{\mu\nu}=\partial_\mu A_\nu-\partial_\nu A_\mu$.

    Spontaneous swirl-symmetry breaking via \(\langle H\rangle = v_\text{sw}\) gives masses
    \begin{align}
        M_W &= \frac{1}{2} g_\text{sw}\, v_\text{sw},
        &
        M_Z &= \frac{M_W}{c_\text{sw}},
        &
        M_A &= 0.
    \end{align}
    The swirl-Higgs potential and kinetic term:
    \begin{align}
        \mathcal{L}^{(2)}_{H}&=\frac{1}{2}(\partial_\mu H)(\partial^\mu H) - V(H),\qquad
        V(H)=\frac{\lambda}{4}\,(H^2 - v_\text{sw}^2)^2,
    \end{align}
    with \(\lambda>0\).  The would-be Goldstones \(\phi^0,\phi^\pm\) become the longitudinal modes of \(Z,W^\pm\) in $R_\xi$ gauges.

    \noindent\textbf{Explicit cubic interactions (from non-Abelian structure):}
    \begin{align}
        \mathcal{L}^{(2)}_{\text{cubic}}
        &= i g_\text{sw}\Big[
            (\partial_\mu W_\nu^+ - \partial_\nu W_\mu^+) W^{-\mu} \mathcal{A}^{3\nu}
            -(\partial_\mu W_\nu^- - \partial_\nu W_\mu^-) W^{+\mu} \mathcal{A}^{3\nu}
            \nonumber\\
            &\qquad
            +(\partial_\mu \mathcal{A}_\nu^3 - \partial_\nu \mathcal{A}_\mu^3) W^{+\mu} W^{-\nu}
            \Big] \;+\; \text{mixing into } (Z_\mu, A_\mu)\text{ via } \theta_\text{sw} \\
        &= i g_\text{sw}\Big\{
        c_\text{sw}\big[
            (\partial_\mu W_\nu^+ - \partial_\nu W_\mu^+) W^{-\mu} Z^{\nu}
            -(\partial_\mu W_\nu^- - \partial_\nu W_\mu^-) W^{+\mu} Z^{\nu}
            +( \partial_\mu Z_\nu - \partial_\nu Z_\mu) W^{+\mu} W^{-\nu}
            \big]\nonumber\\
        &\quad + s_\text{sw}\big[
            (\partial_\mu W_\nu^+ - \partial_\nu W_\mu^+) W^{-\mu} A^{\nu}
            -(\partial_\mu W_\nu^- - \partial_\nu W_\mu^-) W^{+\mu} A^{\nu}
            +( \partial_\mu A_\nu - \partial_\nu A_\mu) W^{+\mu} W^{-\nu}
            \big]
        \Big\}. \label{eq:cubic}
    \end{align}

    \noindent\textbf{Explicit quartic interactions (from \(f^{abc}f^{ade}\) terms):}
    \begin{align}
        \mathcal{L}^{(2)}_{\text{quartic}}
        &= -\frac{g_\text{sw}^2}{2}\Big[
            W_\mu^+ W^{-\mu} W_\nu^+ W^{-\nu}
            - W_\mu^+ W_\nu^- W^{+\mu} W^{-\nu}
            \Big]\nonumber\\
        &\quad - g_\text{sw}^2 \Big[
            c_\text{sw}^2 \big( Z_\mu Z^\mu W_\nu^+ W^{-\nu} - Z_\mu W^{+\mu} Z_\nu W^{-\nu} \big)
            + s_\text{sw}^2 \big( A_\mu A^\mu W_\nu^+ W^{-\nu} - A_\mu W^{+\mu} A_\nu W^{-\nu} \big)\nonumber\\
            &\qquad\qquad
            + s_\text{sw} c_\text{sw} \big( Z_\mu A^\mu W_\nu^+ W^{-\nu} - Z_\mu W^{+\mu} A_\nu W^{-\nu} - A_\mu W^{+\mu} Z_\nu W^{-\nu} + A_\mu Z^\mu W_\nu^+ W^{-\nu}\big)
            \Big]. \label{eq:quartic}
    \end{align}

    \noindent\textbf{Vector--Higgs interactions (mass-generating and higher):}
    \begin{align}
        \mathcal{L}^{(2)}_{\text{VH}}
        &= g_\text{sw} M_W\, H\, W_\mu^+ W^{-\mu}
        + \frac{g_\text{sw} M_Z}{2 c_\text{sw}^2}\, H\, Z_\mu Z^\mu
        \nonumber\\
        &\quad
        - \frac{g_\text{sw}^2}{4}\, H^2 \left(2\,W_\mu^+ W^{-\mu} + \frac{1}{c_\text{sw}^2} Z_\mu Z^\mu\right)
        % Goldstone-gradient terms (\phi^0,\phi^\pm) in R_\xi\ gauges
        + \text{(Goldstone-gradient terms $\phi^0,\phi^\pm$ in $R_\xi$ gauges)}
        \label{eq:VH}
    \end{align}

    All the long SM structures in your block (2) are generated by Eqs.~\eqref{eq:cubic},\eqref{eq:quartic},\eqref{eq:VH} after expanding $s_\text{sw},c_\text{sw}$ and canonical normalizations.

% ===========================================================
% (3) FERMIONIC KNOTS + GAUGE COUPLINGS
% ===========================================================
    \paragraph{(3) Knotted fermions and swirl charges (SM block 3).}
    For each knot species \(K\in\{\ell, \nu, u, d, \dots\}\), with topological charges \((Q_A^K,Q_Z^K,T^\pm)\) fixed by knot/linking and chirality:
    \begin{align}
        \mathcal{L}^{(3)}_{\text{knot-fermions}}
        &= \sum_K \bar{\Psi}_K \left(i\gamma^\mu \partial_\mu - m_K\right)\Psi_K
        + \sum_K \bar{\Psi}_K \gamma^\mu \left( g_A Q_A^K A_\mu + g_Z Q_Z^K Z_\mu \right)\Psi_K
        \nonumber\\
        &\quad + \frac{g_W}{\sqrt{2}} \left[ \bar{\Psi}_u \gamma^\mu T^+ \Psi_d\, W_\mu^+ + \bar{\Psi}_d \gamma^\mu T^- \Psi_u\, W_\mu^- \right],
    \end{align}
    where the physical couplings are
    \begin{align}
        g_W &= g_\text{sw},\qquad
        g_Z = \frac{g_\text{sw}}{c_\text{sw}},\qquad
        g_A = g_\text{sw}\, s_\text{sw}.
    \end{align}
    Masses come from the VAM mass functional adopted in your project:
    \begin{align}
        \boxed{
            m_K \;=\; \frac{4}{\alphaEM\,\phiG}\,
            \Big(\tfrac{1}{2}\,\rhoe\,\Ce^2\Big)\,
            V_K \,\xi(n_K),
            \qquad \xi(n)\equiv 1+\beta\log n,
        }
    \end{align}
    with \(V_K\) the core volume tied to the knot geometry (and \(\rc\)), and \(\beta\) a fitted coherence parameter.

% ===========================================================
% (4) YUKAWA-LIKE COUPLINGS TO SWIRL-HIGGS
% ===========================================================
    \paragraph{(4) Knot--Higgs (density) interactions (SM block 4).}
    In the broken phase,
    \begin{align}
        \mathcal{L}^{(4)}_{\text{Yukawa-like}}
        &= -\sum_K y_K\, H\, \bar{\Psi}_K \Psi_K
        -\sum_K \tilde{y}_K\, \phi^0\, \bar{\Psi}_K i\gamma^5 \Psi_K
        \nonumber\\
        &\quad - \sum_{K,K'} \left[ y_{KK'}^{(+)}\, \phi^+\, \bar{\Psi}_K T^+ \Psi_{K'}
                                 + y_{KK'}^{(-)}\, \phi^-\, \bar{\Psi}_K T^- \Psi_{K'}  \right],
    \end{align}
    with the \emph{geometric Yukawa} identification
    \begin{align}
        \boxed{
            y_K \;=\; \frac{m_K}{v_\text{sw}} \;=\;
            \frac{4}{\alphaEM\,\phiG}\,
            \frac{(\tfrac{1}{2}\rhoe \Ce^2)\, V_K\, \xi(n_K)}{v_\text{sw}}.
        }
    \end{align}
    The $\phi^0,\phi^\pm$ interactions reproduce the axial/charged currents as in SM (your long block 4), now topologically organized.

% ===========================================================
% (5) GHOSTS, GAUGE-FIXING, CONSTRAINTS (incl. X,Y analogs)
% ===========================================================
    \paragraph{(5) Gauge-fixing, FP ghosts, Clebsch ghosts (SM block 5 analog).}
    \begin{align}
        \mathcal{L}^{(5)}_{\text{ghosts/gauge-fix}}
        &= -\frac{1}{2\xi}\,(\partial^\mu \mathcal{A}^a_\mu)^2
        + \bar{c}^a\, \partial^\mu D_\mu^{ab} c^b
        + \bar{\tilde c}\,\partial^2 \tilde c \quad \text{(Clebsch-Abelian)},
    \end{align}
    where \(D_\mu^{ab}=\partial_\mu\delta^{ab}+g_\text{sw} f^{acb}\mathcal{A}_\mu^c\).
    The additional $(X^\pm,X^0,Y)$ of your block (5) can be identified with (linear combinations of) non-physical ghost doublets associated to the $SU(2)$ and Abelian Clebsch sectors, with mass parameters tied to the gauge parameter $\xi$ and $M_W,M_Z$ in $R_\xi$ gauges:
    \begin{align}
        \mathcal{L}^{(5)}_{X,Y}
        &= \bar{X}^+(\partial^2-\xi M_W^2)X^+
        +\bar{X}^-(\partial^2-\xi M_W^2)X^-
        +\bar{X}^0(\partial^2-\xi M_Z^2)X^0
        +\bar{Y}\,\partial^2 Y \nonumber\\
        &\quad + i g_\text{sw}\, c_\text{sw}\,
        \big[ W^{+\mu}(\partial_\mu \bar{X}^0 X^- - \partial_\mu \bar{X}^+ X^0)
        + W^{-\mu}(\partial_\mu \bar{X}^- X^0 - \partial_\mu \bar{X}^0 X^+) \big] \nonumber\\
        &\quad + i g_\text{sw}\, s_\text{sw}\,
        \big[ W^{+\mu}(\partial_\mu \bar{Y} X^- - \partial_\mu \bar{X}^+ Y)
        + W^{-\mu}(\partial_\mu \bar{X}^- Y - \partial_\mu \bar{Y} X^+)\big] \nonumber\\
        &\quad + \dots\quad \text{(further ghost--Goldstone terms per gauge choice)}.
    \end{align}

    \paragraph{Incompressibility and absolute-time constraints.}
    \begin{align}
        \mathcal{L}_{\text{constraints}}
        = \lambda\,(\nabla\!\cdot\!\vec v)
        + \eta\, (u_\mu u^\mu - 1),
    \end{align}
    with Lagrange multipliers \(\lambda,\eta\).

% ===========================================================
% PARAMETER RELATIONS IN TERMS OF USER CONSTANTS
% ===========================================================
    \paragraph{Parameter summary in your constants (ready to plug numerically).}
    \begin{align}
        \kappa_\omega &= \rhof\,\Ce^2,
        &
        g_\text{sw} &= \frac{\Ce}{\rc}\,\Cg,
        &
        v_\text{sw} &= \sqrt{\chi\,\frac{\rhoe}{\Ce^2}},
        \\[4pt]
        M_W &= \frac{1}{2} g_\text{sw} v_\text{sw}
        = \frac{1}{2}\,\frac{\Ce}{\rc}\,\Cg \,\sqrt{\chi\,\frac{\rhoe}{\Ce^2}}
        = \frac{\Cg \sqrt{\chi}}{2\rc}\,\sqrt{\rhoe},
        \\
        M_Z &= \frac{M_W}{c_\text{sw}},\qquad M_A=0, \qquad
        m_K = \frac{4}{\alphaEM\,\phiG}\Big(\tfrac{1}{2}\rhoe \Ce^2\Big)V_K\,\xi(n_K),\qquad
        y_K = \frac{m_K}{v_\text{sw}}.
    \end{align}
    Note the particularly transparent expression \(M_W=\tfrac{\Cg \sqrt{\chi}}{2\rc}\sqrt{\rhoe}\) independent of \(C_e\) after substitution.

% ===========================================================
% LIMIT CHECKS & DIMENSIONAL CONSISTENCY
% ===========================================================
    \paragraph{Sanity checks.}
    \emph{Irrotational limit:} \(\vec\omega\to 0\Rightarrow \mathcal{A}_\mu^a\to 0\), so vectors decouple;
    \(\mathcal{L}\to \tfrac{1}{2}(\partial H)^2 - V(H) + \sum_K \bar{\Psi}_K(i\slashed{\partial}-m_K)\Psi_K\).\\
    \emph{Dimensions:} \([\kappa_\omega]=\text{J m}^{-3}\), \([\mathcal{W}] = \text{s}^{-1}\), so \([\kappa_\omega\,\mathcal{W}^2]=\text{J m}^{-3}\).
    \(g_\text{sw}\) has \([\text{s}^{-1}]\); with $v_\text{sw}$ carrying the scalar-field unit, $M_{W,Z}$ carry energy units (set $c=\hbar=1$ inside the field sector if desired; or keep SI explicitly).

% ===========================================================
% REFERENCES TO VAM PAPERS (BIBTEX KEYS)
% ===========================================================
    \begin{thebibliography}{9}

        \bibitem{Iskandarani2025-VAM0}
        O.~Iskandarani,
        \emph{Revisiting the \AE ther: From Einstein to the Vortex Fluid Paradigm} (VAM-0), Zenodo (2025).
% DOI as in your record

        \bibitem{Iskandarani2025-TimeDilation}
        O.~Iskandarani,
        \emph{Time Dilation in a 3D Superfluid \AE ther Model} (VAM-1), Zenodo (2025).

        \bibitem{Iskandarani2025-SwirlClocks}
        O.~Iskandarani,
        \emph{Swirl Clocks and Vorticity-Induced Gravity} (VAM-2), Zenodo (2025).

        \bibitem{Iskandarani2025-Benchmark}
        O.~Iskandarani,
        \emph{Benchmarking Vortex \AE ther Model vs General Relativity} (VAM-3), Zenodo (2025).

        \bibitem{Iskandarani2025-EmergentGR}
        O.~Iskandarani,
        \emph{Emergent General Relativity from Structured Swirl Dynamics in VAM} (VAM-4), Zenodo (2025).

        \bibitem{Iskandarani2025-KnottedGauge}
        O.~Iskandarani,
        \emph{Knotted Gauge Fields} (VAM-6), Zenodo (2025).

        \bibitem{Iskandarani2025-MasterMass}
        O.~Iskandarani,
        \emph{Master Equation for Particle Masses} (VAM-11), Zenodo (2025).

        \bibitem{Iskandarani2025-TopLag}
        O.~Iskandarani,
        \emph{Topological Fluid Dynamics Lagrangian} (VAM-14), Zenodo (2025).

        \bibitem{Iskandarani2025-QuantumGravity}
        O.~Iskandarani,
        \emph{Quantum Gravity via Superfluid Vorticity} (VAM-15), Zenodo (2025).

    \end{thebibliography}

\end{document}