%! Author = Omar Iskandarani
%! Title = Swirl Clocks and Vorticity-Induced Gravity
%! Date = May 23, 2025
%! Affiliation = Independent Researcher, Groningen, The Netherlands
%! License = CC-BY 4.0
%! ORCID = 0009-0006-1686-3961
%! DOI = 10.5281/zenodo.15566336


\documentclass[a4paper,12pt]{article}

% Page Geometry
\usepackage[a4paper, margin=2cm]{geometry}

% Language, Encoding, Fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[english]{babel}

% Colors, Graphics, Diagrams
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{xcolor}

% Math and Physics
\usepackage{amsmath, amssymb, physics}
\usepackage{siunitx}

% Tables and Figures
\usepackage{float}
\usepackage{booktabs}
\usepackage{array, tabularx, makecell, multirow}
\renewcommand{\arraystretch}{1.5}
\renewcommand{\floatpagefraction}{.8}
\usepackage[font=footnotesize]{caption}
\usepackage{subcaption}

% Code and Listings
\usepackage{listings}
\lstset{basicstyle=\ttfamily\footnotesize, breaklines=true}

% TOC Customization
\usepackage{tocloft}
\setcounter{tocdepth}{4}
\renewcommand{\cftsecfont}{\bfseries}
\renewcommand{\cftsubsecfont}{\itshape}
\renewcommand{\cftsecleader}{\cftdotfill{5}}
\renewcommand{\contentsname}{\centering \Huge\textbf{Contents}}

% Links and Metadata
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue,
    pdftitle={The Vortex Æther Model},
    pdfauthor={Omar Iskandarani},
    pdfkeywords={vorticity, gravity, æther, fluid dynamics, time dilation, VAM}
}
\usepackage{bookmark} % PDF bookmarks

% Bibliography
\usepackage[numbers]{natbib} % Or switch to biblatex if preferred

% Line and Hyphenation
\usepackage[none]{hyphenat}
\usepackage{amsfonts}

\sloppy


\begin{document}


    \author{
        Omar Iskandarani\\
        \small Independent Researcher, Groningen, The Netherlands
        \thanks{\texttt{info@omariskandarani.com}}
        \thanks{ORCID: \href{https://orcid.org/0009-0006-1686-3961}{0009-0006-1686-3961} \quad DOI: \href{https://doi.org/10.5281/zenodo.15566336}{10.5281/zenodo.15566336} \quad License: \href{https://creativecommons.org/licenses/by/4.0/}{CC-BY 4.0}}
        \noindent\thanks{\textbf{Keywords:} \textit{time dilation, superfluid æther, Vortex Æther Model, vortex dynamics, emergent time, fluid spacetime, special relativity, analog gravity, 3D vortex structures, quantized circulation, relativistic effects, topological matter, fluid mechanics, vortex clocks, knot theory, mass generation, swirl gravity, topological quantum field theory, Gross--Pitaevskii, Biot--Savart, Standard Model unification, periodic table topology}}
    }

    \title{
        \textbf{gregergreg}\\[0.5em]
        \large Reformulating ... \\
        \normalsize A Topological ...
    }
    \date{\today}

    \maketitle
    \begin{abstract}
    We propose the Vortex Æther Model (VAM), a fluid-dynamic reformulation of the Standard Model and gravitation based on structured vorticity fields in a Euclidean 3D space with absolute time. In contrast to spacetime curvature approaches, VAM attributes mass, charge, spin, and field interactions to topologically stable vortex knots, rings, and braids embedded in an incompressible, inviscid superfluid æther. We present ten foundational benchmarks that reconstruct particle and field dynamics from first principles:

    \begin{enumerate}
        \item A massless photon emerges as a dipole vortex ring with net translational velocity induced by asymmetric swirl and source-sink pressure gradients.
        \item Electromagnetic fields are reproduced from vortex-induced velocity and circulation patterns, recovering dipole fields and field lines from Biot–Savart analogs.
        \item Energy, circulation, and helicity densities are derived from the vorticity field, yielding a conserved fluid Hamiltonian formulation.
        \item The electron is modeled as a trefoil knot \( T(2,3) \), with topological spin-\( \frac{1}{2} \), negative helicity, and intrinsic charge linked to net swirl.
        \item The proton arises as a triple-link of unknots with chiral handedness, encoding confinement and net positive charge through topological asymmetry.
        \item The neutron is represented by a Borromean ring configuration—globally bound yet locally unlinked—resulting in a neutral but metastable composite structure.
        \item The neutrino is identified with a null-knot configuration possessing zero net helicity and time-sensitive chirality, explaining its weak interaction profile and matter-antimatter asymmetry.
        \item A Planck-scale compact vortex yields an upper bound on the universal force (\( F_\text{\ae}^\text{max} \)), derived from energy density gradients and core swirl velocities, linking gravitational limits to vortex compactness.
        \item The gravitational potential is reconstructed from vorticity-induced swirl fields and compared to the Schwarzschild potential. The VAM potential remains finite at small radii and decays exponentially, resolving singularities inherent in general relativity.
        \item Finally, we reformulate the Standard Model Lagrangian in terms of vortex operators, helicity couplings, and topological invariants. Electric charge emerges from integrated helicity, spin from knot class, and mass from stored Bernoulli energy. Massive bosons and the Higgs arise as bifurcations in swirl field configurations.

    \end{enumerate}

    Together, these benchmarks demonstrate that known physical particles and fields can be recovered from a single coherent framework based on topological vortex dynamics in the æther. The model provides testable predictions, resolves classical divergences, and offers a physically intuitive, mathematically precise foundation for unifying quantum mechanics and gravitation without invoking curved spacetime.
    \end{abstract}

        \input{Sections/SwirlclockInducedTimeAsymmetry}
        \input{Sections/SwirlclockInterference andTimeReversalAsymmetryNeutrinoOscillations}
        \input{Sections/TopologicalOriginOfCharge}
        \input{Sections/GravityAndElectromagnetismUnification}
        \input{Sections/DeriveMaxwellFromLagrangian}



    \appendix
    \input{Appendix/Appendix_PhotonAsDipole}
    \input{Appendix/Appendix_BenchmarkCoulombLaw}
    \input{Appendix/Appendix_MagneticFields}
    \input{Appendix/Appendix_Benchmark3}
    \input{Appendix/Appendix_Benchmark4}
    \input{Appendix/Appendix_Benchmark5}
    \input{Appendix/Appendix_Benchmark6}
    \input{Appendix/Appendix_Benchmark7}
    \input{Appendix/Appendix_Benchmark8}
    \input{Appendix/Appendix_Benchmark9}
    \input{Appendix/Appendix_Benchmark10}


    \bibliographystyle{unsrt}
    \bibliography{../references}



\end{document}
