%! Author = Omar Iskandarani
%! Title = ...
%! Date = xx-xx-2025
%! Affiliation = Independent Researcher, Groningen, The Netherlands
%! License = © 2025 Omar Iskandarani. All rights reserved. This manuscript is made available for academic reading and citation only. No republication, redistribution, or derivative works are permitted without explicit written permission from the author. Contact: info@omariskandarani.com
%! ORCID = 0009-0006-1686-3961
%! DOI = 10.5281/zenodo.xxxxxxx

% === Metadata ===
\newcommand{\papertitle}{Photon as a Topological Vortex Ring: \\ Torsion and the Geometry of Light in the Æther}
\newcommand{\paperdoi}{10.5281/zenodo.xxxxxxxx}

\documentclass[twocolumn,aps,pre,floatfix,nofootinbib]{revtex4-2}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{physics}
\usepackage{caption}
\usepackage{bm}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{mathtools}
\usetikzlibrary{knots,intersections,decorations.pathreplacing}
\usetikzlibrary{3d, calc, arrows.meta, positioning}
\usepackage{pgfmath}
\usetikzlibrary{decorations.pathmorphing}
\pgfplotsset{compat=1.18} % or version you have
\usepackage{titlesec}
\usepackage{ulem}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\renewcommand{\grqq}{``}
\usepackage{subfiles}
\usepackage{ragged2e}
\usepackage{amsmath, amssymb, bm}

\begin{document}
    \title{\papertitle}
    \author{Omar Iskandarani}
    \thanks{Independent Researcher, Groningen, The Netherlands\\
    info@omariskandarani.com \\
    ORCID: \href{https://orcid.org/0009-0006-1686-3961}{0009-0006-1686-3961} \\
    DOI: \href{https://doi.org/\paperdoi}{\paperdoi}}
    \date{\today}




    \begin{abstract}
        \vspace*{-0.5em}
        \section*{\centering Abstract}
        \vspace*{-1em}
    We propose a reformulation of the photon as a quantized, massless vortex ring embedded in an incompressible, inviscid superfluid æther. Using Cartan's geometric structure equations, we draw a formal correspondence between torsional and curvature defects in condensed matter (dislocations and disclinations) and the vortex-core features of ætheric flow. The Burgers and Frank vectors of defect theory find natural analogs in swirl density and angular vorticity. We show how the photon’s Lagrangian, Hamiltonian, and Jacobian structure follow directly from vortex ring energetics in this medium. By coupling this with a Biot–Savart swirl framework and time dilation due to swirl velocity, we reproduce the null proper time of the photon from first principles. This unification provides a geometric and fluid-mechanical basis for the photon's quantized behavior and opens the door for a deeper topological reinterpretation of gauge bosons in the Standard Model.
    \end{abstract}
    \maketitle




    \section{Introduction}
        In the Vortex \ae ther Model (VAM), gravitation and quantum phenomena are reformulated through the topology and dynamics of vorticity in an incompressible, inviscid fluid-like \ae ther. This article derives a formal connection between classical defect theory---dislocations and disclinations in condensed matter physics---and VAM's vortex core structures using the language of Cartan geometry. Inspired by recent developments \cite{kobayashi2025}, we reinterpret torsion and curvature within the Cartan structure equations in terms of swirl density and topological vortex curvature. Furthermore, we show that light itself---the photon---can be interpreted as a topologically stable vortex ring, unifying electromagnetic energy propagation with the mechanics of superfluid circulation.


    \section{Vortex Ring and Cartan Torsion}

    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            % Ring (vortex core)
            \draw[thick, blue!70] (0,0) circle (2);

            % Flow lines (tangential swirl)
            \foreach \a in {15,45,75,105,135,165,195,225,255,285,315,345} {
                \draw[->, red!70, thick]
                ({2*cos(\a)}, {2*sin(\a)})
                ++({-0.4*sin(\a)}, {0.4*cos(\a)})
                -- ++({0.4*sin(\a)}, {-0.4*cos(\a)});
            }

            % Central torsion arrow
            \draw[->, thick, violet] (0,0) -- (0,2.6) node[above] {\small $\mathbf{T}^y$};

            % Axis labels
            \draw[->, thick, gray!60] (0,0) -- (3,0) node[right] {$x$};
            \draw[->, thick, gray!60] (0,0) -- (0,3) node[above] {$y$};

            % Label
            \node at (0,-2.5) {\small Vortex ring with tangential velocity $C_e$};
        \end{tikzpicture}
    \end{center}


    \noindent
    This diagram shows a vortex ring in the $x$-$y$ plane with tangential swirl velocity (red arrows) and central torsion aligned along the $y$-axis,
    representing $\mathbf{T}^y$ from Cartan's torsion 2-form.


    \input{sections/01_Framework}


    \input{sections/02_Biot-Savart}



    \input{sections/ConclusionOutlook}

    \appendix
    \input{sections/A1_Derivation_of_Lagrangian_Hamiltonian_Jacobian}
% ============= END of content ============

    \bibliographystyle{unsrt}
    \bibliography{photon_vortex_refs}

\end{document}