\section{Gauge Interaction Rules and Vortex Dynamics}

To reproduce Standard Model interaction mechanisms in VAM, we reinterpret gauge boson exchanges as topological transformations within the vortex Æther manifold.

\subsection{Topological Transitions as Interactions}

Particle interactions correspond to changes in knot configuration via reconnection, twisting, and braiding operations. These transformations affect linking number, helicity, and orientation:

\begin{itemize}
    \item \textbf{Vortex Reconnection}: Local reconnections shift knot type, interpreted as particle decay or scattering.
    \item \textbf{Twist Induction}: Changes internal helicity, analogous to photon emission or absorption.
    \item \textbf{Braiding Interactions}: Exchange of braid positions changes color charge, modeling gluon exchange.
\end{itemize}

\subsection{Dynamical Equations of Vorticity Transfer}

We generalize the Navier–Stokes vorticity equation to encode gauge-like dynamics:
\begin{equation}
    \partial_t \vec{\omega} + (\vec{v} \cdot \nabla)\vec{\omega} = (\vec{\omega} \cdot \nabla)\vec{v} + \nu \nabla^2 \vec{\omega} + \vec{F}_{\text{gauge}},
\end{equation}
where $\vec{F}_{\text{gauge}}$ encodes reconnection-based topological force analogues of gauge field interactions.

\subsection{Reconnection Operators as Gauge Vertices}

Gauge boson exchanges are modeled using helicity-rotation operators $\mathcal{R}^{(a)}$ acting on knot wavefunctions:
\begin{equation}
    D_\mu \psi_K = \partial_\mu \psi_K + i g \mathcal{R}^{(a)}_\mu \psi_K,
\end{equation}
analogous to covariant derivatives in quantum field theory. The index $a$ labels the generator (color, isospin, etc.) of the transformation.

These operators obey an algebra:
\begin{equation}
[\mathcal{R}^{(a)}_\mu, \mathcal{R}^{(b)}_\nu] = i f^{abc} \mathcal{R}^{(c)}_\lambda,
\end{equation}
mirroring the Lie algebra structure of $SU(3)$ and $SU(2)$ gauge theories.

\subsection{Decay and Scattering Channels}

Using topological conservation laws (helicity, linking number, writhe), allowed and forbidden decay pathways can be predicted:
\begin{itemize}
    \item \textbf{Conservation of Helicity:} $H_{\text{in}} = H_{\text{out}}$
    \item \textbf{Knot-Symmetry Selection Rules:} Only symmetric-to-symmetric or chirality-preserving reconnections are energetically favorable.
\end{itemize}

\textbf{Example:} A doubly linked torus knot representing a muon may decay into a single trefoil (electron) and an untwisting spiral (neutrino), conserving linking number.