\section{Boundary and Gauge Conditions in VAM}

To ensure physical consistency, topological conservation, and a well-posed variational principle in the Vortex Æther Model (VAM), appropriate boundary and gauge conditions must be imposed on all dynamical fields. These conditions guarantee finite energy configurations, preserve topological structure, and define allowable transformations analogous to gauge freedom in field theory.

\subsection{Boundary Conditions}

The vortex and scalar fields in VAM are localized structures embedded in a compressible æther background. The following boundary conditions ensure that solutions are physically acceptable:

\begin{align*}
    \vec{v}(\vec{x}, t) &\rightarrow 0 \quad \text{as} \quad |\vec{x}| \rightarrow \infty && \text{(vanishing velocity)} \\
    \rho(\vec{x}, t) &\rightarrow \rho_0 = \text{const.} && \text{(uniform background density)} \\
    \phi(\vec{x}, t) &\rightarrow \phi_{\text{vac}} && \text{(vacuum scalar potential)} \\
    \vec{\omega}(\vec{x}, t) &= \nabla \times \vec{v} \rightarrow 0 && \text{(localized vorticity)} \\
    \int \vec{v} \cdot \vec{\omega} \, d^3x &< \infty && \text{(finite helicity integral)}
\end{align*}

Additionally, knotted vortex configurations must be closed, non-self-intersecting, and topologically quantized to ensure particle-like stability and mass conservation.

\subsection{Gauge Conditions}

Although VAM does not contain gauge fields in the traditional sense, several fluid-dynamic symmetries mirror the structure of gauge theories in the Standard Model. These \grqq fluid gauges\textquotedblright can be expressed as follows:

\begin{enumerate}
    \item \textbf{Velocity Potential Gauge (Irrotational Decomposition):}
    \[
        \vec{v} = \nabla \psi + \nabla \times \vec{A}
    \]
    where $\psi$ is the scalar velocity potential and $\vec{A}$ is a swirl vector potential. The system is invariant under the transformation $\vec{A} \to \vec{A} + \nabla \chi$.

    \item \textbf{Incompressibility Constraint (Coulomb Gauge Analog):}
    \[
        \nabla \cdot \vec{v} = 0
    \]
    which corresponds to a divergence-free æther flow, consistent with a near-incompressible medium and fluid analogs of gauge fixing.

    \item \textbf{Topological Gauge Invariance:}
    The identity of vortex particles is encoded in their knot topology (e.g., trefoil, figure-eight). Gauge transformations must preserve topological invariants such as linking number and helicity:
    \[
        \mathcal{H} = \int \vec{v} \cdot \vec{\omega} \, d^3x = \text{constant}
    \]
    These invariants act as topological charges analogous to electric or color charge.
\end{enumerate}

These boundary and gauge conditions collectively constrain the solution space of the VAM Lagrangian and ensure consistency with observed quantum behavior, mass conservation, and particle stability.
