%! Author = Omar Iskandarani
%! Date = 2/15/2025
\documentclass[a4paper,10pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{array}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{physics}
\usepackage{cite}


\geometry{margin=1in}


\title{Exploring the unification of atomic and gravitational phenomena under the Vortex \AE ther Model (VAM)}
\author{Omar Iskandarani}
\date{\today}

\begin{document}
    \maketitle

    \maketitle

    \begin{abstract}

This collection of articles explores the unification of atomic and gravitational phenomena under the Vortex \AE ther Model (VAM), providing a novel interpretation of quantum mechanics and gravity through vorticity dynamics. We propose that both hydrogenic wavefunctions and gravitational fields in VAM are governed by exponentially decaying vortex structures, establishing a deep correspondence between atomic orbitals and large-scale gravitational interactions. Unlike conventional approaches relying on curved spacetime or intrinsic mass, VAM suggests that mass, energy, and time emerge from structured \AE theric vorticity, reinforcing a purely three-dimensional Euclidean framework.

The unification is demonstrated by analyzing electron orbitals as stable vortex states that follow hydrodynamic decay laws, analogous to the gravitational vorticity decay governing massive objects. We derive fundamental relations between the Bohr radius and the Coulomb barrier in VAM, showing that mass accumulation and probability density functions share identical mathematical structures. Through the mapping of vortex quantization principles across scales, we establish that gravitational interactions are the macroscopic limit of the same vortex dynamics that define atomic energy states.

This work further examines implications such as mass-energy equivalence as an emergent property of vorticity, black holes as large-scale quantum vortex states, and time dilation as a direct consequence of vortex confinement rather than a separate spacetime dimension. The conclusions challenge conventional models of particle physics and cosmology, suggesting that all known interactions may ultimately be reformulated within a vorticity-based framework. Experimental validation is proposed through comparisons with superfluid analogs and controlled vortex confinement studies, aiming to solidify VAM as a testable and predictive theory of fundamental physics.

    \end{abstract}
    \input{../src/250_eqs}

%    \newpage
    \input{movement_of_free_Æther_particles}
 %   \newpage
    \input{vortex_pressure}
 %   \newpage
    \input{vorticity_natural_coords}
 %   \newpage\
    \input{300_vorticity_derivation}

 %   \newpage
    \input{relative_vorticity_2knots}


 %   \newpage
    \input{why_coulomb_differs_in_electromagnetic_system}
 %   \newpage
    \input{Spin_Torsion_in_Vortices}
 %   \newpage
    \input{photon_vortex}

 %   \newpage
    \input{atomic_orbits_Vortex_Gravity_space_and_time}
 %   \newpage
    \input{Bridging_Knot_Theory_Particle_Physics}
 %   \newpage
    \input{Casimir_Effect}
 %   \newpage
    \input{BlackBody_radiation_entropy}
 %   \newpage
 %   \newpage
    \input{Æther_gas}

    \input{maxwell}
 %   \newpage
    \input{vortex_energy}
 %   \newpage
    \input{sprites}

 %   \newpage
    \input{oumuamua}








    \bibliographystyle{ieeetr}
    \bibliography{../src/references}

\end{document}