
\section{Electromagnetism: Twist and Chirality}

In VAM, the familiar forces of electricity and magnetism are no longer mysterious fields floating in space—they are special patterns of swirl and twist in the cosmic æther. This approach makes electromagnetism almost as tangible as watching waves form on a pond or feeling the wind twist around your body.


\subsection*{Charge as Twist: The Vortex Signature}

Imagine every knot in the æther (every particle) is like a spinning whirlpool. The way it twists—whether clockwise or counterclockwise—determines its “charge.”


\begin{itemize}

\item
A vortex knot spinning one way (say, right-handed) has a positive charge.




\item
The same knot, spun in the opposite direction (left-handed), has a negative charge.




\end{itemize}

So, just as two whirlpools swirling the same way might join forces, but those spinning in opposite ways push each other apart, the twist (chirality) sets the rules for how particles attract or repel.


\subsection*{Electromagnetic Waves as Coordinated Swirls}

Electromagnetic waves—like light—are not abstract ripples, but waves of twist and coordinated motion in the æther. Picture many tiny whirlpools popping in and out of existence, their twists aligned so that a pulse (a photon) travels smoothly across the pond. The dance of twist is what carries light and energy from place to place.


\subsection*{Chirality: The Handedness of Nature}

Chirality isn’t just about left or right. In VAM, it’s the core property that gives each knot its unique electromagnetic “personality.” If you look at your left and right hands, you can’t rotate one to become the other—they are mirror images. In the same way, every knot’s twist gives it a fixed identity, setting its behavior in electromagnetic interactions.


\subsection*{Analogy Bite}

\begin{itemize}

\item
“Electromagnetic forces are like the push and pull of neighboring whirlpools—the direction they spin decides if they combine, clash, or send out ripples of light.”




\end{itemize}