\section{Swirl-Induced Time and Clockwork in Vortex Knots}

In the Vortex Æther Model (VAM), stable knots are understood not just as matter structures, but as the fundamental agents of time. Their internal swirl—tangential rotation with speed \( C_e \) around a core radius \( r_c \)—generates an asymmetric stress field in the surrounding æther. This asymmetry produces a persistent \textbf{axial flow along the knot core}, functionally equivalent to a local "time-thread". Despite the absence of a literal helical geometry, the knot behaves as if it winds the æther forward—\textit{acting as a screw-like conductor of time}.

\subsection*{Cosmic Swirl Orientation}

Like magnetic domains, vortex knots exhibit a preferred swirl direction. In a universe with a dominant chirality, reversing a knot's rotation (as in antimatter) may only yield a stable structure in isolation. This offers a natural explanation for:
\begin{itemize}
    \item the scarcity of antimatter in the visible universe,
    \item the unidirectionality of macroscopic time,
    \item and the alignment of clock rates across vast cosmic domains.
\end{itemize}

\subsection*{Swirl as a Local Time Carrier}

The central swirl generates a flow \( \vec{v}_\text{time} \) which sets the local pace of time evolution:
\[
    dt_{\text{local}} \propto \frac{dr}{\vec{v} \cdot \vec{\omega}}
\]
Time is thus not fundamental in VAM—it is emergent from the flow alignment between velocity \( \vec{v} \) and vorticity \( \vec{\omega} \). The product \( \vec{v} \cdot \vec{\omega} \) measures helicity flux, and its inverse governs the rate at which a system evolves.

\subsection*{Networks of Temporal Flow}

Vortex knots tend to organize along extended swirl lines—akin to iron filings aligning along magnetic field lines. Around mass concentrations, these lines form structured bundles of time flow, which naturally explain:
\begin{itemize}
    \item gravitational attraction as a gradient of swirl density,
    \item local time dilation near massive bodies,
    \item the global arrow of time as a circulation topology in the æther.
\end{itemize}

This emergent swirl-clock mechanism unifies mass, time, and directionality into a single fluid-geometric framework. It replaces spacetime curvature with conserved helicity flow, offering a fully mechanical origin for temporal dynamics.
