\section{Foundational Fields and Geometry}
\paragraph{Effective densities (mainstream field-theory style).}
\begin{itemize}
    \item \rhoF \equiv \text{effective fluid density}, [-0.8]
    \item \rhoE \equiv \tfrac12\,\rhoF\,\vnorm^{2}\quad(\text{swirl energy density}),
    \item \rhoM \equiv \rhoE/c^{2}\quad(\text{mass-equivalent density}).
\end{itemize}


We work on a 4D Lorentzian manifold with metric \(g_{\mu\nu}\) of signature \((-+++)\). The totally antisymmetric tensor satisfies \(\varepsilon^{0123}=+1\); antisymmetrization is \(X_{[\mu\nu]}=\tfrac12(X_{\mu\nu}-X_{\nu\mu})\). Indices are raised/lowered with \(g_{\mu\nu}\).

\begin{itemize}
 \item \textbf{Clock Field \(T(x)\) and Preferred Foliation.}
 Define the unit timelike condensate 4-velocity
 \begin{equation}
 u_{\mu} \equiv \frac{\partial_{\mu}T}{\sqrt{-\,\partial_{\alpha}T\, \partial^{\alpha}T}},
 \qquad u_\mu u^\mu=-1,
 \end{equation}
 which is invariant under monotone reparametrizations \(T\!\to\! f(T)\).
 The spatial projector onto leaves orthogonal to \(u_\mu\) is
    \begin{equation}
        h_{\mu\nu} \equiv g_{\mu\nu}+u_\mu u_\nu, \qquad h_{\mu\nu}u^\nu=0, \qquad h^\mu{}_\alpha h^\alpha{}_\nu=h^\mu{}_\nu.
    \end{equation}
    Integrability of the foliation is equivalent to the Frobenius condition \(u_{[\mu}\nabla_{\nu}u_{\rho]}=0\), which holds when \(u_\mu\propto\partial_\mu T\).

    \item \textbf{Condensate Modulus \(\Phi\).}
    A real scalar controlling medium scales (stiffness, characteristic speeds). We write \(\Phi(x)=\Phi_0+\delta\Phi(x)\) about a homogeneous vacuum value \(\Phi_0>0\). This \emph{is not} the Standard Model Higgs; it does not introduce Yukawa parameters.

    \item \textbf{Two-Form Potential \(B_{\mu\nu}\) and Three-Form Field Strength \(H\).}
    \begin{equation}
        H_{\mu\nu\rho} \equiv \partial_{[\mu} B_{\nu\rho]} \quad (\;= \tfrac{1}{3!}\,\varepsilon_{\mu\nu\rho\sigma}\,\tilde H^{\sigma}\;).
    \end{equation}
    Gauge symmetry: \(B \mapsto B + d\Lambda\) with 1-form \(\Lambda_\mu\). Bianchi identity: \(\partial_{[\sigma}H_{\mu\nu\rho]}=0\).
    Swirl strings couple electrically to \(B\) (worldsheet term \(\int_\Sigma B\)); their topological charge is measured by fluxes of \(H\).

    \item \textbf{Emergent Swirl Connection \(\mathcal{W}_\mu^a\).}
    A non-Abelian gauge potential for coarse-grained vorticity modes valued in a compact Lie algebra with structure constants \(f^{abc}\).
    \begin{equation}
        \mathcal{W}_{\mu\nu}^a \equiv \partial_\mu \mathcal{W}_\nu^a - \partial_\nu \mathcal{W}_\mu^a + g_{\!sw}\, f^{abc}\,\mathcal{W}_\mu^b \mathcal{W}_\nu^c,
        \qquad \tilde{\mathcal{W}}^{\mu\nu a} \equiv \tfrac12 \varepsilon^{\mu\nu\rho\sigma}\mathcal{W}_{\rho\sigma}^a .
    \end{equation}
    Gauge transformations act as \(\delta \mathcal{W}_\mu^a = -(\nabla_\mu \alpha)^a - g_{\!sw} f^{abc}\alpha^b \mathcal{W}_\mu^c\),
    and on matter via the covariant derivative \(D_\mu=\nabla_\mu + g_{\!sw}\,\mathcal{W}_\mu^a T^a\) with generators \(T^a\).

    \item \textbf{Knot Fermion Fields \(\Psi_K\).}
    Effective relativistic spinors associated with stable knotted swirl strings labeled by topological class \(K\) (e.g., torus knots). Their rest masses are non-perturbative soliton energies \(m_K^{(\mathrm{sol})}\). They transform under the swirl gauge group via \(D_\mu\Psi_K\).

\end{itemize}

\paragraph{Dimensional assignments (natural units \(\hbar=c=1\)).}
\([ \mathcal{W}_\mu ]=1,\; [ \mathcal{W}_{\mu\nu} ]=2,\; [ B_{\mu\nu} ]=1,\; [ H_{\mu\nu\rho} ]=2,\; [ \Phi ]=1,\; [ \Psi ]=\tfrac{3}{2},\; [g_{\!sw}]=0.\)

% --- Integrated: Clock Foliation Dynamics and Constraints ---
\subsection{Clock Foliation Dynamics and Constraints}
Define the unit timelike foliation vector by
\begin{equation}
u_\mu \equiv \frac{\partial_\mu T}{\sqrt{-g^{\alpha\beta}\partial_\alpha T\partial_\beta T}},\qquad u_\mu u^\mu=-1.
\end{equation}
We include the foliation-unit (khronon) sector
\begin{align}
\mathcal L_{T} \;=\; \frac{M_u^{2}}{2}\Big[&c_1(\nabla_\mu u_\nu)(\nabla^\mu u^\nu)+c_2(\nabla_\mu u^\mu)^2+c_3(\nabla_\mu u_\nu)(\nabla^\nu u^\mu)\\
&+c_4\,u^\mu u^\nu (\nabla_\mu u_\alpha)(\nabla_\nu u^\alpha)\Big]
\;+\;\lambda\,(u_\mu u^\mu+1).\nonumber
\end{align}
This sector is invariant under monotone reparametrizations \(T\mapsto f(T)\), and since \(u\propto\nabla T\) the foliation is integrable (Frobenius).

A cosmological origin follows from a shift--symmetric condensate \(T(x)=\mu t+\pi(x)\) with timelike gradient \(X=g^{\mu\nu}\partial_\mu T\partial_\nu T<0\) and
\(\mathcal L_{\rm khr}(X)=M_\star^4\,P(X/M_\star^4)\). In unitary gauge this reproduces \(\mathcal L_T\).

\paragraph{Observational constraints.}
The tensor-mode speed is
\(c_T^2=\frac{1}{1-c_{13}}\) with \(c_{13}=c_1+c_3\). The binary neutron-star event GW170817/GRB170817A implies \(|c_T/c-1|\lesssim10^{-15}\), so we take the baseline fit \(c_{13}=0\).
Preferred-frame PPN parameters \(\alpha_1(c_i),\alpha_2(c_i)\) constrain the remaining combinations, and the absence of gravitational \v{C}erenkov losses requires \(c_T\ge 1\).
Non-gravitational LIV constraints can be tracked via the SME Data Tables.

\begin{table}[h]
\centering
\caption{Clock-sector parameters and baseline constraint.}
\begin{tabular}{lcc}
\hline
Parameter & Meaning & Baseline / Constraint \\
\hline
$c_{13} \equiv c_1 + c_3$ & controls tensor speed $c_T^2=\frac{1}{1-c_{13}}$ & set to $0$ (GW170817) \\
$c_1,c_2,c_4$ & foliation-vector couplings in $\mathcal{L}_T$ & to be bounded in scans \\
$M_u$ & foliation sector scale & free EFT scale \\
\hline
\end{tabular}
\end{table}

\paragraph{Coarse--graining between leaf rotation and fluid density.}
\begin{equation}
K \equiv \frac{\rhocore\, r_c}{\vswirl},
\qquad
\rhoF \;=\; K\,\Omega .
\end{equation}
With Canon constants: \(K = 5.01509060\times 10^{-3}\ \mathrm{kg\,m^{-3}\,s}\), \(\Omega^\ast = 1.39578735\times 10^{-4}\ \mathrm{s^{-1}}\), and \(T^\ast=2\pi/\Omega^\ast\approx 12.50\ \mathrm{h}\).
% --- End Integrated block ---
