%! Author = Omar Iskandarani
%! Date = 2025-06-01
%! Title = Glossary of Vorticity, Helicity, and Rotational Flow Terms

\documentclass[12pt]{article}
\usepackage[a4paper,margin=2cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{physics}
\title{Glossary of Vorticity, Helicity, and Rotational Flow Terms}
\author{Omar Iskandarani}
\date{June 2025}

\begin{document}
\maketitle

\section*{Introduction}
This glossary summarizes key concepts in rotational fluid dynamics, both in classical fluid mechanics and within the framework of the Vortex \AE{}ther Model (VAM). Each term is defined with relevant physical context and mathematical expressions.

\section*{Glossary}




\section*{Local Rotation Concepts (Directly related to Vorticity)}

\textbf{Angular Velocity Vector \( \boldsymbol{\Omega} \)} \\
Describes the instantaneous rotation rate of a fluid element. \\
Related to vorticity by:
\[
\boldsymbol{\omega} = 2 \boldsymbol{\Omega}
\]
\textit{Role:} Connects fluid kinematics to rigid-body rotation analogy.

\medskip
\textbf{Curl of Velocity \( \boldsymbol{\omega} = \nabla \times \vec{v} \)} \\
This is the formal definition of vorticity. \\
\textit{Tensorial Form:} Antisymmetric part of the velocity gradient.

\section*{Twisting and Linking (Helicity-related Quantities)}

\textbf{Kinetic Helicity \( H = \int \vec{v} \cdot \boldsymbol{\omega} \, dV \)} \\
Scalar, conserved in ideal flows (incompressible, inviscid). \\
Measures alignment between flow velocity and vorticity. \\
Positive/Negative values indicate right/left-handed twist.

\medskip
\textbf{Magnetic Helicity \( H_m = \int \vec{A} \cdot \vec{B} \, dV \)} \\
Analogue of kinetic helicity for magnetohydrodynamics (MHD). \\
Where \( \vec{B} = \nabla \times \vec{A} \) is the magnetic field.

\medskip
\textbf{Cross Helicity \( H_c = \int \vec{v} \cdot \vec{B} \, dV \)} \\
Measures alignment of velocity and magnetic field in MHD. \\
Shows dynamic coupling between flow and magnetic field lines.

\medskip
\textbf{Relative Helicity} \\
Normalized form of helicity to compare twistedness independent of flow strength:
\[
H_{\text{rel}} = \frac{\int \vec{v} \cdot \boldsymbol{\omega} \, dV}
{\left( \int |\vec{v}|^2 dV \right)^{1/2}
\left( \int |\boldsymbol{\omega}|^2 dV \right)^{1/2}}
\]

\section*{Flow Structure Types (Swirl \& Vorticity-related)}

\textbf{Spiral Vortex / Swirl Flow} \\
Flow pattern with both radial and tangential components. \\
Described often in cylindrical coordinates as:
\[
\vec{v}(r,\theta,z) = v_r(r,z) \, \hat{r} + v_\theta(r,z) \, \hat{\theta} + v_z(r,z) \, \hat{z}
\]
Appears in tornadoes, hurricanes, and swirl combustors.

\medskip
\textbf{Rankine Vortex} \\
A model vortex with solid-body rotation inside a core and irrotational flow outside:
\[
\omega(r) =
\begin{cases}
\text{constant}, & r < r_c \\
0, & r > r_c
\end{cases}
\]

\medskip
\textbf{Burgers Vortex} \\
A steady solution to Navier-Stokes that balances vorticity diffusion and stretching:
\[
\omega(r) = \omega_0 e^{-a r^2}
\]
Useful in modeling turbulence and vortex stretching.

\medskip
\textbf{Vortex Sheet} \\
Surface with discontinuous tangential velocity but continuous normal component. \\
Idealization of shear layers.

\medskip
\textbf{Vortex Filament / Line} \\
Line-like structure where vorticity is concentrated (mathematically a delta function). \\
Central to Biot–Savart law and Kelvin's circulation theorem.

\section*{Derived Quantities \& Diagnostics}

\textbf{Q-Criterion} \\
Identifies vortices by comparing strain and rotation:
\[
Q = \frac{1}{2} \left( \| \boldsymbol{\Omega} \|^2 - \| \mathbf{S} \|^2 \right)
\]
where \( \mathbf{S} \) is the rate-of-strain tensor and \( \boldsymbol{\Omega} \) is the rotation tensor.

\medskip
\textbf{Vorticity Magnitude \( |\boldsymbol{\omega}| \)} \\
Scalar field used to visualize rotation intensity.

\medskip
\textbf{Enstrophy \( E = \frac{1}{2} \int |\boldsymbol{\omega}|^2 \, dV \)} \\
Analogous to energy, but for rotation. \\
Important in 2D turbulence where it is approximately conserved.

\medskip
\textbf{Circulation \( \Gamma = \oint_{\mathcal{C}} \vec{v} \cdot d\vec{l} \)} \\
Line integral of velocity around a closed curve. \\
Related to vorticity via Stokes' theorem:
\[
\Gamma = \iint_S \boldsymbol{\omega} \cdot \hat{n} \, dA
\]

\section*{In VAM-Specific Context}

\textbf{Swirl Clock Time Rate} \\
Local time evolution determined by:
\[
d\tau = dt \sqrt{1 - \frac{|\boldsymbol{\omega}|^2}{C_e^2}}
\]
Corresponds to time dilation via rotation in VAM.

\medskip
\textbf{Swirl Lagrangian} \\
Includes helicity or knottedness:
\[
\mathcal{L}_{\text{swirl}} = \lambda \vec{v} \cdot (\nabla \times \vec{v}) = \lambda \vec{v} \cdot \boldsymbol{\omega}
\]

\medskip
\textbf{Topological Charge / Linking Number} \\
Quantifies the entanglement of vortex tubes:
\[
H = \sum_i \Gamma_i^2 \, \text{Lk}_i
\]
Related to the Hopf invariant in field theory.


\subsection*{Vorticity $\boldsymbol{\omega}$}
The curl of the velocity field:
\[ \boldsymbol{\omega} = \nabla \times \mathbf{v} \]
It describes the local spinning motion of a fluid parcel. In VAM, vorticity modulates local time and governs pressure differentials.

\subsection*{Relative Vorticity}
Vorticity measured relative to a rotating frame. Important in geophysical flows and rotating reference systems in æther dynamics.

\subsection*{Angular Velocity Vector $\boldsymbol{\Omega}$}
Describes the solid-body rotation rate of a fluid element, related to vorticity via:
\[ \boldsymbol{\omega} = 2\boldsymbol{\Omega} \]

\subsection*{Vortex Tube}
A bundle of vortex lines forming a tubular region of concentrated vorticity. Vortex tubes can evolve into knotted structures in VAM.

\subsection*{Vortex Line}
A curve everywhere tangent to the vorticity vector field. These lines illustrate the direction and connectivity of local rotational motion.

\subsection*{Vortex Stretching}
Occurs when a vortex tube is elongated by the flow. In 3D, this process increases the magnitude of vorticity, conserving angular momentum.

\subsection*{Helicity $H$}
A scalar defined as:
\[ H = \int \mathbf{v} \cdot \boldsymbol{\omega} \, dV \]
It measures the extent to which velocity and vorticity are aligned. It is a topological invariant in ideal fluids and central to vortex knot theory in VAM.

\subsection*{Kinetic Helicity Density}
The pointwise scalar product:
\[ h = \mathbf{v} \cdot \boldsymbol{\omega} \]
Gives the local density of helicity, whose integral over a volume is total helicity.

\subsection*{Cross Helicity}
Used in magnetohydrodynamics (MHD):
\[ H_c = \int \mathbf{v} \cdot \mathbf{B} \, dV \]
Measures the alignment between fluid velocity and magnetic field.

\subsection*{Magnetic Helicity}
Another MHD quantity:
\[ H_m = \int \mathbf{A} \cdot \mathbf{B} \, dV \]
where $\mathbf{B} = \nabla \times \mathbf{A}$. It quantifies the linkage of magnetic field lines.

\subsection*{Swirl}
A qualitative term indicating rotational flow. In cylindrical coordinates, the azimuthal component $v_\theta$ is often referred to as the swirl velocity.

\subsection*{Swirl Clock}
In VAM, a clock governed by the internal angular frequency $\omega_0$ of a vortex core. Time dilation is given by:
\[ d\tau = dt \sqrt{1 - \frac{\omega^2}{C_e^2}} \]

\subsection*{Swirl Lagrangian}
A field Lagrangian incorporating helicity terms:
\[ \mathcal{L}_{\text{swirl}} = \lambda (\mathbf{v} \cdot \boldsymbol{\omega}) \]
This encapsulates the topological and dynamical behavior of swirl fields in VAM.

\subsection*{Rotation}
A general term describing angular motion in fluids, encompassing both local (vorticity) and global (rigid body) rotations.

\subsection*{Circulation $\Gamma$}
Defined by:
\[ \Gamma = \oint_{\mathcal{C}} \mathbf{v} \cdot d\mathbf{l} \]
Measures the total rotation around a closed loop. Stokes’ theorem connects it to surface-integrated vorticity.

\subsection*{Enstrophy $\mathcal{E}$}
A quadratic measure of rotational intensity:
\[ \mathcal{E} = \frac{1}{2} \int |\boldsymbol{\omega}|^2 \, dV \]
Important in turbulence analysis and vortex conservation.

\subsection*{Q-Criterion}
A scalar diagnostic for vortex identification:
\[ Q = \frac{1}{2}(||\boldsymbol{\Omega}||^2 - ||\mathbf{S}||^2) \]
where $\mathbf{S}$ is the strain-rate tensor. Positive $Q$ indicates vortex-dominated regions.

\subsection*{Rankine Vortex}
A model vortex with solid-body rotation inside a core ($r < r_c$) and irrotational flow outside ($r > r_c$). Simplified but useful for teaching.
\[ \omega(r) =
\begin{cases}
\text{constant}, & r < r_c \\
0, & r > r_c
\end{cases}
\]

\subsection*{Burgers Vortex}
A steady-state vortex model balancing stretching and viscous diffusion:
\[ \omega(r) = \omega_0 e^{-a r^2} \]
Frequently used in studies of turbulent vortex dynamics.

\subsection*{Topological Charge}
Quantifies the entanglement or linkage of vortex lines. In VAM, this relates to conserved quantities such as helicity and can be described using the Hopf invariant.

\section*{References}
\begin{itemize}[leftmargin=1.5em]
    \item G.K. Batchelor, \textit{An Introduction to Fluid Dynamics}, Cambridge University Press, 1967.
    \item H.K. Moffatt, "The degree of knottedness of tangled vortex lines", \textit{J. Fluid Mech.}, vol. 35, 1969, pp. 117–129. doi:10.1017/S0022112069000991
    \item H.K. Moffatt and A. Tsinober, "Helicity and singular structures in fluid dynamics", \textit{Ann. Rev. Fluid Mech.}, vol. 24, 1992, pp. 281–312. doi:10.1146/annurev.fl.24.010192.001433
\end{itemize}

\end{document}
