%! Author = Omar Iskandarani
%! Date = 2/15/2025
\documentclass[a4paper,10pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{array}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{physics}
\usepackage{cite}
\usepackage{natbib}
\usepackage{authblk}
\usepackage{subfiles}
\usepackage{subcaption}
\usepackage{float}
\usepackage{markdown}

\geometry{margin=1in}


\title{The Vortex \AE ther Model: A Vorticity-Based Framework for Gravity and Electromagnetism}
\author{Omar Iskandarani}
\date{\today}

\begin{document}
    \maketitle

    \maketitle

    \begin{abstract}
        This paper introduces the Vortex \AE ther Model (VAM), a novel approach to fundamental physics where gravity and electromagnetism emerge from vorticity fields in an incompressible, inviscid \AE ther.
        The model presented herein offers a modern interpretation of what is conventionally referred to as \AE ther theory, reimagined as a structured, inviscid superfluid medium governed by vorticity interactions rather than classical particulate motion.
        While the 19th-century concept of a luminiferous \AE ther was rejected following the Michelson-Morley experiment, the fundamental questions it sought to address—concerning the nature of space, energy propagation, and fundamental interactions—remain open.
        This work argues that a contemporary \AE theric model, grounded in fluid dynamics and topological vortex structures, may provide novel insights into quantum mechanics, inertia, and gravity.
        Unlike General Relativity, which relies on spacetime curvature, VAM posits that gravitational attraction arises from pressure gradients induced by vortex filaments.
        Electromagnetism, in turn, is described as a consequence of structured vortex networks, with magnetic fields emerging from circulating \AE ther flows.
        Experimental predictions include measurable frequency shifts in rotating Bose-Einstein condensates and anomalous electromagnetic effects in high-vorticity plasmas.
        Proposed laboratory tests and detection methods are outlined to validate the Vortex \AE ther Model.
        By extending Clausius’s thermodynamic principles into a vorticity-based gravitational model, this framework establishes a connection between classical thermodynamics, quantum mechanics, and fluid dynamics.
        Notably:        - Thermal expansion-contraction cycles of vortex knots mirror the behaviors observed in gas expansion laws.
        - Energy transfer within the \AE ther follows structured vorticity dynamics, rather than being mediated by mass-energy interactions.
        - Entropy-driven expansion aligns with cosmological models describing universal inflation without requiring dark energy.
        This model offers a novel perspective on the nature of space, energy, and fundamental interactions, providing a coherent framework for future research into the unification of physical forces.
    \end{abstract}

    \input{00_introduction}

    \section{Part I}\label{sec:part-1}
    \input{01_extension_physics}
    \input{02_observations_on_gr}
    \input{03_æther_5D}
    \input{04_observation_on_light_particle}
    \input{05_riged_body}
    \input{06_derevation_swirl}

    \section{Part II}\label{sec:part-2}
    \input{07_VAM-reformulation-GR}
    \input{08_path-Integral-gauge-relativity}
    \input{09_Magnetism_as_vorticity}



    \input{10_vortex_induced_magnetic_fields}
    \input{11_addressing_qed_corrections}
    \input{12_vam_derevation}
    \input{14_bragg}


%    \input{20_rest}
%    \input{21_vorticity_derivation}

    \bibliographystyle{ieeetr}
    \bibliography{references}

\end{document}